---
title: "Untitled"
output: html_document
date: "2026-03-23"
---

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(ggplot2)
library(reshape2)
library(tidyr)
install.packages("beepr")     # Install beepr package
library(beepr)                # Load beepr package
require("lgpr")
require("rstan")
require("ggplot2")

#install.packages('GauPro')
#library(GauPro)
required <- c("tidyverse", "lubridate", "scales", "patchwork", "viridis")
new      <- required[!required %in% installed.packages()[, "Package"]]
if (length(new)) install.packages(new, repos = "https://cloud.r-project.org")
 
suppressPackageStartupMessages({
  library(tidyverse)
  library(lubridate)
  library(scales)
  library(patchwork)
  library(viridis)
})

set.seed(123)
```


```{r}
#data <- read_csv( "C:/Users/zubareva/OneDrive - Tilburg University/Desktop/code/phd/data examples/ESMdata/ESMdata.csv")
data <- read_csv( "C:/Users/margo/OneDrive/Рабочий стол/phd code/data examples/ESMdata/ESMdata.csv")

names(data)[names(data)=="...1"] <- "datapoint"
#names<- colnames(data)
#mood_names <- names[grepl( "mood_", names, fixed = TRUE)]
#mood_data<-data[c("datapoint",mood_names)]
#mood_data<-data[c("datapoint", "mood_satisfi")]

#mdf <- melt(mood_data[1:500,],id.vars="datapoint")
#ggplot(mdf, aes( x=datapoint, y=value, colour=variable, group=variable )) + geom_line()

df <- data %>%
  mutate(
    # --- temporal parsing ---
    date_parsed = dmy(date),
    hour        = hour(hms(beeptime)),
    tired        = as.integer(phy_tired),#  
    irritated = as.integer(mood_irritat), #mood_satisfi, mood_enthus, mood_down, mood_lonely, 
 
    # --- derived time components ---
    datetime = as.POSIXct(as.character(paste(date_parsed, beeptime)), format="%Y-%m-%d %H:%M:%S"),
    timestamp = as.numeric(datetime),# the number of seconds that have passed between date object and 1/1/1970
    year      = year(date_parsed),
    month_num = month(date_parsed),
    date_num = day(date_parsed),
    month     = month(date_parsed, label = TRUE, abbr = TRUE),
    week      = isoweek(date_parsed),
    wday_num  = wday(date_parsed, week_start = 1),
    wday      = wday(date_parsed, label = TRUE, abbr = TRUE, week_start = 1)
  )  %>%
  filter(!is.na(date_parsed), tired >= 1, tired <= 7, irritated >= 1, irritated <= 7)
 
```

```{r}
# one row per calendar day
daily <- df %>%
  group_by(date_parsed, year, month, month_num, week, wday, wday_num) %>%
  summarise(
    tired_mean = mean(tired),
    tired_sd   = sd(tired),
    irr_mean = mean(irritated),
    irr_sd = sd(irritated),
   

    n_beeps   = n(),
    .groups   = "drop"
  )

daily <- daily %>% mutate(
  timestamp = as.numeric(date_parsed),
  day = row_number(),
  day_of_year = as.numeric(strftime(date_parsed, format = "%j")),
  day_of_month = day(date_parsed)
)
 
df_dep<- df %>% 
  filter(!is.na(dep))

# weekly depression scores
weekly <- df_dep %>%
  group_by(year, week) %>%
  summarise(
    dep_mean  = mean(dep),
    .groups    = "drop"
  )

# one row per calendar month
monthly <- df %>%
  group_by(year, month_num, month) %>%
  summarise(
    tired_mean = mean(tired),
    tired_se   = sd(tired) / sqrt(n()),
    irr_mean = mean(irritated),
    irr_se = sd(irritated) / sqrt(n()),
    n         = n(),
    .groups   = "drop"
  ) %>%
  mutate(date_label = as.Date(paste(year, month_num, "01", sep = "-")))
 
# profile by hour of day
hourly <- df %>%
  group_by(hour) %>%
  summarise(
    tired_mean = mean(tired),
    tired_se   = sd(tired) / sqrt(n()),
    irr_mean = mean(irritated),
    irr_se = sd(irritated) / sqrt(n()),
    n         = n(),
    .groups   = "drop"
  )
 
# profile by weekday
wday_agg <- df %>%
  group_by(wday, wday_num) %>%
  summarise(
    tired_mean = mean(tired),
    tired_se   = sd(tired) / sqrt(n()),
    irr_mean = mean(irritated),
    irr_se = sd(irritated) / sqrt(n()),
    n         = n(),
    .groups   = "drop"
  ) %>%
  arrange(wday_num)

mday_agg <- df %>%
  group_by(date_num) %>%
  summarise(
    tired_mean = mean(tired),
    tired_se   = sd(tired) / sqrt(n()),
    irr_mean = mean(irritated),
    irr_se = sd(irritated) / sqrt(n()),
    n         = n(),
    .groups   = "drop"
  ) %>%
  arrange(date_num)
 
weekly_d <- ggplot(weekly, aes(seq(1, 28, by = 1), dep_mean)) +
  geom_point() +  geom_smooth(
    method = "loess", span = .1, se = TRUE,
    color = clr_line, fill = clr_line, alpha = .15, linewidth = .9
  )

```
#------ plots
```{r}
clr_bg     <- "#FFFFFF" #"#0f1117"
clr_panel  <- "#FFFFFF"#"#181b24"
clr_grid   <- "#262b38"
clr_text   <- "#c8cfe0"
clr_sub    <- "#6b7694"
clr_line   <- "#7eb8f7"
clr_accent <- "#f97316"
clr_pos    <- "#4ade80"
clr_mid    <- "#f0c040"
clr_neg    <- "#fb7185"
 
mood_gradient <- scale_color_gradient2(
  low = clr_neg, mid = clr_mid, high = clr_pos,
  midpoint = 3.5, guide = "none"
)
mood_fill_gradient <- scale_fill_gradient2(
  low = clr_neg, mid = clr_mid, high = clr_pos,
  midpoint = 3.5, guide = "none"
)
 
theme_mood <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.background  = element_rect(fill = clr_bg,    color = NA),
      panel.background = element_rect(fill = clr_panel, color = NA),
      panel.grid.major = element_line(color = clr_grid, linewidth = .3),
      panel.grid.minor = element_blank(),
      axis.text        = element_text(color = clr_sub,  size = rel(.85)),
      axis.title       = element_text(color = clr_text, size = rel(.9)),
      plot.title       = element_text(color = clr_text, size = rel(1.15),
                                      face = "bold", margin = margin(b = 3)),
      plot.subtitle    = element_text(color = clr_sub,  size = rel(.82),
                                      margin = margin(b = 8)),
      legend.background = element_rect(fill = clr_panel, color = NA),
      legend.text       = element_text(color = clr_sub),
      legend.title      = element_text(color = clr_text),
      strip.text        = element_text(color = clr_text, face = "bold"),
      strip.background  = element_rect(fill = clr_grid,  color = NA),
      plot.margin       = margin(12, 14, 10, 14)
    )
}
theme_set(theme_mood())
```

```{r}
grand_mean <- mean(df$tired, na.rm = TRUE)
# important: overall
## 4a. Overall trajectory — raw daily means + LOESS ───────────────────────────
p_overall <- ggplot(daily, aes(date_parsed, tired_mean)) +
  geom_point(aes(color = tired_mean, size = n_beeps)) +
  geom_smooth(
    method = "loess", span = .1, se = TRUE,
    color = clr_line, fill = clr_line, alpha = .15, linewidth = .9
  ) +
  geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
  mood_gradient +
  scale_y_continuous(limits = c(1, 6), breaks = 1:6) +
    scale_size_continuous(range = c(.5, 2.5), name = "N beeps") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  labs(
    title    = "Overall Mood Trajectory",
    subtitle = "Daily mean · LOESS smooth (span = 0.10) · dashed = grand mean",
    x = NULL, y = "Mood satisfaction (1–6)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
 
 ## 4c. Monthly line + SE ribbon ───────────────────────────────────────────────
p_monthly <- ggplot(monthly, aes(date_label, tired_mean)) +
  geom_ribbon(
    aes(ymin = tired_mean - tired_se, ymax = tired_mean + tired_se),
    fill = clr_line, alpha = .2
  ) +
  geom_line(color = clr_line, linewidth = .9) +
  geom_point(aes(size = n, color = tired_mean)) +
  mood_gradient +
  geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
  scale_size_continuous(range = c(1.5, 5), name = "N beeps") +
  scale_y_continuous(limits = c(1, 6), breaks = 1:6) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b\n%Y") +
  labs(
    title    = "Monthly Mood Averages",
    subtitle = "Line ± 1 SE ribbon · point size = number of beeps",
    x = NULL, y = "Mean mood"
  ) +
  theme(legend.position = "right")
 
# important: weekly
## 4e. Day-of-week bar chart ───────────────────────────────────────────────────
p_wday <- ggplot(wday_agg, aes(wday_num, tired_mean)) +
    geom_ribbon(
    aes(ymin = tired_mean - tired_se, ymax = tired_mean + tired_se),
    fill = clr_line, alpha = .2
  ) +
  geom_line(color = clr_line, linewidth = .9)+
  # geom_errorbar(
  #   aes(ymin = tired_mean - tired_se, ymax = tired_mean + tired_se),
  #   width = .2, color = clr_text, linewidth = .5
  # )+
    geom_point(aes(color = tired_mean, size = n)) +
  mood_gradient +
  geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
    scale_size_continuous(range = c(1.5, 5), name = "N") +
  scale_x_continuous(
    breaks = seq(1, 7, 1),
    labels = paste0(c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
  ) +
  scale_y_continuous(limits = c(0, 6), breaks = 1:6) +
  labs(
    title    = "Mood by Day of Week",
    subtitle = "Mean ± 1 SE · dashed = grand mean",
    x = NULL, y = "Mean mood"
  )

p_mday <- ggplot(mday_agg, aes(date_num, tired_mean)) +
    geom_ribbon(
    aes(ymin = tired_mean - tired_se, ymax = tired_mean + tired_se),
    fill = clr_line, alpha = .2
  ) +
  geom_line(color = clr_line, linewidth = .9)+
  # geom_errorbar(
  #   aes(ymin = tired_mean - tired_se, ymax = tired_mean + tired_se),
  #   width = .2, color = clr_text, linewidth = .5
  # )+
    geom_point(aes(color = tired_mean, size = n)) +
  mood_gradient +
  geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
    scale_size_continuous(range = c(1.5, 5), name = "N") +
  scale_x_continuous(
    breaks = seq(1, 31, 1),
    labels = paste0(c(seq(1, 31, 1)))
  ) +
  scale_y_continuous(limits = c(0, 6), breaks = 1:6) +
  labs(
    title    = "Mood by Day of Month",
    subtitle = "Mean ± 1 SE · dashed = grand mean",
    x = NULL, y = "Mean mood"
  )

 # important: hourly
## 4f. Hour-of-day profile ────────────────────────────────────────────────────
p_hour <- ggplot(hourly, aes(hour, tired_mean)) +
  geom_ribbon(
    aes(ymin = tired_mean - tired_se, ymax = tired_mean + tired_se),
    fill = clr_line, alpha = .2
  ) +
  geom_line(color = clr_line, linewidth = .9) +
  geom_point(aes(color = tired_mean, size = n)) +
  mood_gradient +
    geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
    scale_size_continuous(range = c(1.5, 5), name = "N") +
  scale_x_continuous(
    breaks = seq(6, 23, 1),
    labels = paste0(seq(6, 23, 1), ":00")
  ) +
  scale_y_continuous(limits = c(1, 6), breaks = 1:6) +
  labs(
    title    = "Mood by Hour of Day",
    subtitle = "Mean ± 1 SE across all beeps at that hour",
    x = "Hour", y = "Mean mood"
  )

p_overall
p_monthly
 p_wday
p_hour
p_mday

```

```{r}
grand_mean <- mean(df$irritated, na.rm = TRUE)
# important: overall
## 4a. Overall trajectory — raw daily means + LOESS ───────────────────────────
p_overall <- ggplot(daily, aes(date_parsed, irr_mean)) +
  geom_point(aes(color = irr_mean, size = n_beeps)) +
  geom_smooth(
    method = "loess", span = .1, se = TRUE,
    color = clr_line, fill = clr_line, alpha = .15, linewidth = .9
  ) +
  geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
  mood_gradient +
  scale_y_continuous(limits = c(1, 6), breaks = 1:6) +
    scale_size_continuous(range = c(.5, 2.5), name = "N beeps") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  labs(
    title    = "Overall Mood Trajectory",
    subtitle = "Daily mean · LOESS smooth (span = 0.10) · dashed = grand mean",
    x = NULL, y = "Mood satisfaction (1–6)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
 


 ## 4c. Monthly line + SE ribbon ───────────────────────────────────────────────
p_monthly <- ggplot(monthly, aes(date_label, irr_mean)) +
  geom_ribbon(
    aes(ymin = irr_mean - irr_se, ymax = irr_mean + irr_se),
    fill = clr_line, alpha = .2
  ) +
  geom_line(color = clr_line, linewidth = .9) +
  geom_point(aes(size = n, color = irr_mean)) +
  mood_gradient +
  geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
  scale_size_continuous(range = c(1.5, 5), name = "N beeps") +
  scale_y_continuous(limits = c(1, 6), breaks = 1:6) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b\n%Y") +
  labs(
    title    = "Monthly Mood Averages",
    subtitle = "Line ± 1 SE ribbon · point size = number of beeps",
    x = NULL, y = "Mean mood"
  ) +
  theme(legend.position = "right")
 

# important: weekly
## 4e. Day-of-week bar chart ───────────────────────────────────────────────────
p_wday <- ggplot(wday_agg, aes(wday_num, irr_mean)) +
    geom_ribbon(
    aes(ymin = irr_mean - irr_se, ymax = irr_mean + irr_se),
    fill = clr_line, alpha = .2
  ) +
  geom_line(color = clr_line, linewidth = .9)+
  # geom_errorbar(
  #   aes(ymin = irr_mean - irr_se, ymax = irr_mean + irr_se),
  #   width = .2, color = clr_text, linewidth = .5
  # )+
    geom_point(aes(color = irr_mean, size = n)) +
  mood_gradient +
  geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
    scale_size_continuous(range = c(1.5, 5), name = "N") +
  scale_x_continuous(
    breaks = seq(1, 7, 1),
    labels = paste0(c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
  ) +
  scale_y_continuous(limits = c(0, 6), breaks = 1:6) +
  labs(
    title    = "Mood by Day of Week",
    subtitle = "Mean ± 1 SE · dashed = grand mean",
    x = NULL, y = "Mean mood"
  )
 

p_mday <- ggplot(mday_agg, aes(date_num, irr_mean)) +
    geom_ribbon(
    aes(ymin = irr_mean - irr_se, ymax = irr_mean + irr_se),
    fill = clr_line, alpha = .2
  ) +
  geom_line(color = clr_line, linewidth = .9)+
  # geom_errorbar(
  #   aes(ymin = irr_mean - irr_se, ymax = irr_mean + irr_se),
  #   width = .2, color = clr_text, linewidth = .5
  # )+
    geom_point(aes(color = irr_mean, size = n)) +
  mood_gradient +
  geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
    scale_size_continuous(range = c(1.5, 5), name = "N") +
  scale_x_continuous(
    breaks = seq(1, 31, 1),
    labels = paste0(c(seq(1, 31, 1)))
  ) +
  scale_y_continuous(limits = c(0, 6), breaks = 1:6) +
  labs(
    title    = "Mood by Day of Month",
    subtitle = "Mean ± 1 SE · dashed = grand mean",
    x = NULL, y = "Mean mood"
  )

 
 # important: hourly
## 4f. Hour-of-day profile ────────────────────────────────────────────────────
p_hour <- ggplot(hourly, aes(hour, irr_mean)) +
  geom_ribbon(
    aes(ymin = irr_mean - irr_se, ymax = irr_mean + irr_se),
    fill = clr_line, alpha = .2
  ) +
  geom_line(color = clr_line, linewidth = .9) +
  geom_point(aes(color = irr_mean, size = n)) +
  mood_gradient +
    geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
    scale_size_continuous(range = c(1.5, 5), name = "N") +
  scale_x_continuous(
    breaks = seq(6, 23, 1),
    labels = paste0(seq(6, 23, 1), ":00")
  ) +
  scale_y_continuous(limits = c(1, 6), breaks = 1:6) +
  labs(
    title    = "Mood by Hour of Day",
    subtitle = "Mean ± 1 SE across all beeps at that hour",
    x = "Hour", y = "Mean mood"
  )

p_overall
p_monthly
 p_wday
p_hour
p_mday
```


```{r}
# unimportant
## 4b. Weekly bar chart + LOESS ───────────────────────────────────────────────
p_weekly <- ggplot(weekly, aes(week_start, tired_mean)) +
  geom_col(aes(fill = tired_mean), width = 5, alpha = .85) +
  geom_smooth(
    method = "loess", span = .2, se = FALSE,
    color = clr_line, linewidth = 1
  ) +
  scale_fill_gradient2(
    low = clr_neg, mid = clr_mid, high = clr_pos,
    midpoint = 3.5, name = "Mood"
  ) +
  scale_y_continuous(limits = c(0, 6), breaks = 0:6) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  labs(
    title    = "Weekly Mood Averages",
    subtitle = "Each bar = one ISO week · LOESS trend overlay",
    x = NULL, y = "Mean mood"
  ) +
  theme(
    axis.text.x      = element_text(angle = 30, hjust = 1),
    legend.position  = "right"
  )
 p_weekly

## 4d. Year-over-year seasonal overlay ───────────────────────────────────────
p_yoy <- ggplot(monthly,
                aes(month_num, tired_mean,
                    color = factor(year), group = year)) +
  geom_ribbon(
    aes(ymin = tired_mean - tired_se, ymax = tired_mean + tired_se),
    fill = clr_line, alpha = .2
  )+
  geom_line(linewidth = .9, alpha = .85) +
  geom_point(size = 2) +
  scale_color_viridis_d(option = "C", begin = .15, end = .9, name = "Year") +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  scale_y_continuous(limits = c(1, 6), breaks = 1:6) +
  labs(
    title    = "Year-over-Year Seasonal Pattern",
    subtitle = "Monthly mean per year overlaid for comparison",
    x = NULL, y = "Mean mood"
  ) +
  theme(legend.position = "right")
 p_yoy
 



## 4g. Calendar heatmap — week × weekday, faceted by year ─────────────────────
heat_data <- daily %>%
  group_by(year, week, wday, wday_num) %>%
  summarise(tired_mean = mean(tired_mean), .groups = "drop")
 
p_heat <- ggplot(heat_data,
                 aes(x = week, y = fct_rev(wday), fill = tired_mean)) +
  geom_tile(color = clr_bg, linewidth = .3) +
  facet_wrap(~year, ncol = 1) +
  scale_fill_gradient2(
    low = clr_neg, mid = clr_mid, high = clr_pos,
    midpoint = 3.5, na.value = clr_panel,
    name = "Mood", limits = c(1, 6)
  ) +
  scale_x_continuous(
    breaks = c(1, 13, 26, 39, 52),
    labels = c("Jan", "Apr", "Jul", "Oct", "Dec")
  ) +
  labs(
    title    = "Calendar Heatmap — Daily Mood",
    subtitle = "Each tile = one day (mean of all beeps that day)",
    x = "Week of year", y = NULL
  ) +
  theme(
    panel.grid   = element_blank(),
    axis.text.y  = element_text(size = 8),
    legend.position = "right"
  )

## 4h. Distribution violin + boxplot by year ──────────────────────────────────
p_dist <- df %>%
  mutate(year = factor(year)) %>%
  ggplot(aes(year, tired, fill = year)) +
  geom_violin(alpha = .55, trim = FALSE, linewidth = .3) +
  geom_boxplot(width = .12, outlier.shape = NA,
               color = clr_text, alpha = .7) +
  scale_fill_viridis_d(option = "C", begin = .15, end = .9, guide = "none") +
  scale_y_continuous(breaks = 1:6) +
  labs(
    title    = "Mood Distribution by Year",
    subtitle = "Violin + box · each point = one beep",
    x = NULL, y = "Mood satisfaction (1–6)"
  )
```

```{r}
# ── 5.  Compose pages with patchwork ─────────────────────────────────────────
ann_theme <- theme(
  plot.background = element_rect(fill = clr_bg, color = NA),
  plot.title      = element_text(color = clr_text, size = 16,
                                 face = "bold", margin = margin(b = 4)),
  plot.subtitle   = element_text(color = clr_sub,  size = 11,
                                 margin = margin(b = 6))
)
 
## Page 1 — longitudinal overview
page1 <- (p_overall / p_weekly / p_monthly) +
  plot_annotation(
    title    = "Mood Over Time — Overview",
    subtitle = "Beep-level data aggregated to daily → weekly → monthly",
    theme    = ann_theme
  )
 
## Page 2 — cyclical patterns
page2 <- (p_yoy | p_dist) / (p_wday | p_hour) +
  plot_annotation(
    title    = "Mood Patterns — Seasonal, Weekly & Daily Rhythms",
    theme    = ann_theme
  )
 
## Page 3 — calendar heatmap (height scales with number of years)
n_years <- n_distinct(df$year)
page3   <- p_heat +
  plot_annotation(
    title = "Calendar Heatmap — Daily Mood Satisfaction",
    theme = ann_theme
  )
 

# ── 6.  Save ─────────────────────────────────────────────────────────────────

ggsave("overall_tir.png", p_overall, width = 10, height = 6,
        dpi = 180)
ggsave("month_tir.png", p_monthly, width = 10, height = 6,
        dpi = 180)
ggsave("wday_tir.png", p_wday, width = 10, height = 6,
        dpi = 180)
ggsave("hour_tir.png", p_hour, width = 10, height = 6,
        dpi = 180)
# ggsave("mood_overview.png",  page1, width = 14, height = 16,
#        dpi = 180, bg = clr_bg)
# ggsave("mood_patterns.png",  page2, width = 16, height = 12,
#        dpi = 180, bg = clr_bg)
# ggsave("mood_heatmap.png",   page3, width = 16,
#        height = pmax(4, n_years * 4) + 3,
#        dpi = 180, bg = clr_bg)
 
message("Saved: mood_overview.png  |  mood_patterns.png  |  mood_heatmap.png")
```

#---- df shorter plots
```{r}
df_shorter<- df[c("timestamp", "mood_irritat", "mood_satisfi", "mood_enthus", "mood_down", "mood_lonely", "phy_tired")]
df_shorter
```

```{r}
set.seed(123)

ggplot() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(x = "Time", y = "Irritation") +
  theme(legend.position = "none") +
  theme(aspect.ratio=0.9) +
geom_smooth(data = df_shorter, aes(x=timestamp, y=mood_irritat)) +
  ggtitle("Irritation levels") +
  theme(plot.title = element_text(size = 13, face = "bold")) 
```


```{r}
plot1 <-     ggplot(df_shorter[1:100,], aes(x=timestamp, y=mood_irritat)) + geom_point()
plot2 <-     ggplot(df_shorter, aes(x=timestamp, y=mood_satisfi)) +geom_point()
plot3 <-     ggplot(df_shorter, aes(x=timestamp, y=mood_lonely)) +geom_point()
plot4 <-     ggplot(df_shorter, aes(x=timestamp, y=phy_tired)) +geom_point()
# Arrange plots in a 2x2 grid
grid.arrange(plot1, plot2, plot3, plot4, ncol = 2)

df_shorter %>%
    gather(key,value, mood_satisfi, mood_irritat, mood_enthus, mood_down, mood_lonely, phy_tired) %>%
    ggplot(aes(value, colour=key)) +
  geom_density(alpha = 0.1) 

df_shorter %>%
    gather(key,value, mood_satisfi, mood_irritat, mood_enthus, mood_down, mood_lonely, phy_tired) %>%
    ggplot(aes(x=timestamp, y=value, colour=key)) +
    geom_point()

df_shorter %>%
    gather(key,value, mood_satisfi, mood_irritat, mood_enthus, mood_down, mood_lonely, phy_tired) %>%
    ggplot(aes(x=timestamp, y=value, colour=key)) +
    geom_line()

df_shorter %>%
    gather(key,value, mood_satisfi, mood_irritat, mood_enthus, mood_down, mood_lonely, phy_tired) %>%
    ggplot(aes(x=timestamp, y=value, colour=key)) +
    geom_smooth()
```
# ---- gau pro analysis

```{r}

ggsave("gp14.png", gpp, width = 10, height = 6,
        dpi = 180)


df_sh <- df_shorter[1:1400,]
gp <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp1<-plot(gp)

gp2 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp2<-plot(gp2)

gp3 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp3<-plot(gp3)

gp4 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp4<-plot(gp4)

gp5 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp5<-plot(gp5)

gp6 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp6<-plot(gp6)

gp7 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp7<-plot(gp7)

gp8 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp8<-plot(gp8)

gp9 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp<-plot(gp9)

gp10 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp<-plot(gp10)

gp11 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp<-plot(gp11)

gp12 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp<-plot(gp12)

gp13 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp<-plot(gp13)

gp14 <- gpkm(df_sh$timestamp, df_sh$phy_tired, kernel = k_Periodic(D=1))
gpp<-plot(gp14)
```

```{r}
summary(gp)

```
```{r}
summary(gp14)

```
```{r}
p_overall <- ggplot(daily, aes(timestamp, mood_mean)) +
  geom_point(aes(color = tired_mean, size = n_beeps)) +
  geom_smooth(
    method = "loess", span = .1, se = TRUE,
    color = clr_line, fill = clr_line, alpha = .15, linewidth = .9
  ) +
  geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
  mood_gradient +
  scale_y_continuous(limits = c(1, 6), breaks = 1:6) +
    scale_size_continuous(range = c(.5, 2.5), name = "N beeps") +
 # scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  labs(
    title    = "Overall Mood Trajectory",
    subtitle = "Daily mean · LOESS smooth (span = 0.10) · dashed = grand mean",
    x = NULL, y = "Mood satisfaction (1–6)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
 

p_overall
```


```{r}
daily_sh<- daily[907:1471,]
p_overall <- ggplot(daily_sh, aes(timestamp, tired_mean)) +
  geom_point(aes(color = tired_mean, size = n_beeps)) +
  geom_smooth(
    method = "loess", span = .1, se = TRUE,
    color = clr_line, fill = clr_line, alpha = .15, linewidth = .9
  ) +
  geom_hline(
    yintercept = grand_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
  mood_gradient +
  scale_y_continuous(limits = c(1, 6), breaks = 1:6) +
    scale_size_continuous(range = c(.5, 2.5), name = "N beeps") +
 # scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  labs(
    title    = "Overall Mood Trajectory",
    subtitle = "Daily mean · LOESS smooth (span = 0.10) · dashed = grand mean",
    x = NULL, y = "Mood satisfaction (1–6)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
 

p_overall
```

```{r}
summary(daily)
```
```{r}
kernel_global<-k_Matern52(beta = 2)

gp_daily_01 <- gpkm(daily_sh$timestamp, daily_sh$tired_mean, kernel = kernel_global)
plot(gp_daily_01)
summary(gp_daily_01)
```


```{r}
kernel_global<-k_Gaussian(beta = 1)

gp_daily_0 <- gpkm(daily_sh$timestamp, daily_sh$tired_mean, kernel = kernel_global)
gpp_d0<-plot(gp_daily_0)
gpp_d0
summary(gp_daily_0)
#ggsave("gp_daily_1.png", gpp_d1, width = 10, height = 6, dpi = 180)
```


```{r}
kernel_daily <- k_Periodic(p  = 7)
kernel_global<-k_Gaussian(beta = 2)

gp_daily_1 <- gpkm(daily_sh$timestamp, daily_sh$tired_mean, kernel = kernel_global + kernel_daily)
gpp_d1<-plot(gp_daily_1)
ggsave("gp_daily_1.png", gpp_d1, width = 10, height = 6,
        dpi = 180)
summary(gp_daily_1)
```

```{r}
kernel_daily <- k_Periodic(p  = 7, alpha = 7)

gp_daily_11 <- gpkm(daily_sh$timestamp, daily_sh$tired_mean, kernel =  kernel_daily)
plot(gp_daily_11)
#ggsave("gp_daily_1.png", gpp_d1, width = 10, height = 6, dpi = 180)
summary(gp_daily_11)
```

```{r}
set.seed(12)
daily_shsh<-daily[1:14,]
#kernel_daily <- k_Periodic(p  = 1.5)

f1 <- gpkm(daily_shsh$timestamp, daily_shsh$tired_mean, kernel =  k_Periodic(p = 0.001))
f10 <- gpkm(daily_shsh$timestamp, daily_shsh$tired_mean, kernel =  k_Periodic(p = 0.14))
f2 <- gpkm(daily_shsh$timestamp, daily_shsh$tired_mean, kernel =  k_Periodic(p = 1))
f3 <- gpkm(daily_shsh$timestamp, daily_shsh$tired_mean, kernel =  k_Periodic(p = 1.75))
f4 <- gpkm(daily_shsh$timestamp, daily_shsh$tired_mean, kernel =  k_Periodic(p = 2))
f5 <- gpkm(daily_shsh$timestamp, daily_shsh$tired_mean, kernel =  k_Periodic(p = 2.5))
f6 <- gpkm(daily_shsh$timestamp, daily_shsh$tired_mean, kernel =  k_Periodic(p = 3))
f7 <- gpkm(daily_shsh$timestamp, daily_shsh$tired_mean, kernel =  k_Periodic(p = 3.5))
f8 <- gpkm(daily_shsh$timestamp, daily_shsh$tired_mean, kernel =  k_Periodic(p = 7))
f9 <- gpkm(daily_shsh$timestamp, daily_shsh$tired_mean, kernel =  k_Periodic(p = 14))
#ggsave("gp_daily_1.png", gpp_d1, width = 10, height = 6, dpi = 180)
#summary(gp_daily_11)

plot(f1) 
plot(f10) 
plot(f2) 
plot(f3) 
plot(f4) 
plot(f5) 
plot(f6) 
plot(f7) 
plot(f8) 
plot(f9)

f9$plotkernel()
```

# ---- what the hell is up with these kernels

```{r}
x<- c(1,  11, 15, 50)
y <- c(1,3, 2, 1)

f <- gpkm(x, y, kernel =  k_Periodic(p = 3.14/5))
f$plotkernel()
plot(f)
```
# ------------------- lgpr

```{r}
prior <- list(
  alpha = normal(mu = 0, sigma = 1), # gaussian for magnitudes <alpha>
  ell = igam(shape = 5, scale = 5) # inverse gamma for lengthscales <ell>
)
```

```{r}
fit <- lgp(irr_mean ~ gp(timestamp), data = daily,#, prior = prior
  iter     = 500,
  chains   = 2,
  refresh  = 50
)
print(fit)
```

```{r}
dat<- data.frame("time"=daily$timestamp, "mean"=daily$irr_mean, "pred"=(c((gp_p@y_mean)[1,])))
dat
mean_n<- mean(dat$mean)
gp_p<- pred(fit)
ggplot(dat, aes(time, pred))+geom_line(color = clr_pos)+ geom_point(aes(x = time, y =mean)) + geom_hline(yintercept =mean_n, color = clr_accent, linetype = "dashed")
```
```{r}
 plot_pred(fit,  group_by = NA, t_name = "timestamp")


plot_components(fit,
 group_by = NA, t_name = "timestamp"
)
```
```{r}
fit <- lgp(irr_mean ~ gp_ns(timestamp), data = daily,#, prior = prior
  iter     = 500,
  chains   = 2,
  refresh  = 50
)
print(fit)
```
# nonstationary kernel produces flat result
```{r}
l<-pred(fit)
 plot_pred(fit,  group_by = NA, t_name = "timestamp")


plot_components(fit,
 group_by = NA, t_name = "timestamp"
)
```

#___________________________________________________STAN_____________________________________________________________________________________

```{r}
# install.packages("tictoc")
# install.packages("cmdstanr")
# install.packages("tinytable")
# install.packages("rprojroot")
library(rprojroot)
root <- has_file(".Bayesian-Workflow-root")$make_fix_file()
library(tidyverse)
library(tictoc)
mytoc <- \() {
  toc(func.toc = \(tic, toc, msg) {
    sprintf("%s took %s sec", msg, as.character(signif(toc - tic, 2)))
  })}
library(cmdstanr)
# CmdStanR output directory makes Quarto cache to work
dir.create(root("stan", "stan_output"), showWarnings = FALSE)
options(cmdstanr_output_dir = root("stan", "stan_output"))
library(posterior)
options(pillar.neg = FALSE,
        pillar.subtle = FALSE,
        pillar.sigfig = 2)
library(tinytable)
options(tinytable_format_num_fmt = "significant_cell",
        tinytable_format_digits = 2,
        tinytable_tt_digits = 2)
library(loo)
library(bayesplot)
theme_set(bayesplot::theme_default(base_family = "sans"))
library(ggdist)
library(patchwork)
library(ggrepel)
library(RColorBrewer)
set1 <- brewer.pal(7, "Set1")
Sys.setlocale("LC_TIME", "en_GB.utf8")
source(root("stan", "plot_helpers.R"))
```

```{r}
irr_g_mean <- mean(df_shorter$mood_irritat)
ti_g_mean <- mean(df_shorter$phy_tired)

irr<-ggplot(df_shorter, aes(timestamp, mood_irritat)) +
  geom_point()+#aes(color = tired_mean, size = n_beeps)) +
  geom_smooth(
    method = "loess", span = .1, se = TRUE,
    color = clr_line, fill = clr_line, alpha = .15, linewidth = .9
  ) +
  geom_hline(
    yintercept = irr_g_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
 # mood_gradient +
  scale_y_continuous(limits = c(1, 7), breaks = 1:7) +
    #scale_size_continuous(range = c(.5, 2.5), name = "N beeps") +
  #scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  labs(
    title    = "Overall Mood Trajectory",
    subtitle = "LOESS smooth (span = 0.10) · dashed = grand mean",
    x = NULL, y = "Mood irritation (1–7)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

ti<-ggplot(df_shorter, aes(timestamp, phy_tired)) +
  geom_point()+#aes(color = tired_mean, size = n_beeps)) +
  geom_smooth(
    method = "loess", span = .1, se = TRUE,
    color = clr_line, fill = clr_line, alpha = .15, linewidth = .9
  ) +
  geom_hline(
    yintercept = ti_g_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
 # mood_gradient +
  scale_y_continuous(limits = c(1, 7), breaks = 1:7) +
    #scale_size_continuous(range = c(.5, 2.5), name = "N beeps") +
  #scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  labs(
    title    = "Overall Mood Trajectory",
    subtitle = "LOESS smooth (span = 0.10) · dashed = grand mean",
    x = NULL, y = "Mood tiredness (1–7)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

irr
ti

ggsave("irr_all_weeks.png", irr, width = 10, height = 6,
        dpi = 180)
ggsave("ti_all_weeks.png", ti, width = 10, height = 6,
        dpi = 180)
```
```{r}
irr<- ggplot(daily, aes(timestamp, irr_mean)) +
  geom_point(aes(color = wday_num, size = n_beeps)) +
  geom_smooth(
    method = "loess", span = .1, se = TRUE,
    color = clr_line, fill = clr_line, alpha = .15, linewidth = .9
  ) +
  geom_hline(
    yintercept = irr_g_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
 # mood_gradient +
  scale_y_continuous(limits = c(1, 7), breaks = 1:7) +
    scale_size_continuous(range = c(.5, 2.5), name = "N beeps") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  labs(
    title    = "Overall Mood Trajectory",
    subtitle = "Daily mean · LOESS smooth (span = 0.10) · dashed = grand mean",
    x = NULL, y = "Mood irritation (1–7)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

ti<- ggplot(daily, aes(timestamp, tired_mean)) +
  geom_point(aes(color = wday_num, size = n_beeps)) +
  geom_smooth(
    method = "loess", span = .1, se = TRUE,
    color = clr_line, fill = clr_line, alpha = .15, linewidth = .9
  ) +
  geom_hline(
    yintercept = ti_g_mean,
    linetype = "dashed", color = clr_accent, linewidth = .5
  ) +
 # mood_gradient +
  scale_y_continuous(limits = c(1, 7), breaks = 1:7) +
    scale_size_continuous(range = c(.5, 2.5), name = "N beeps") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b '%y") +
  labs(
    title    = "Overall Mood Trajectory",
    subtitle = "Daily mean · LOESS smooth (span = 0.10) · dashed = grand mean",
    x = NULL, y = "Mood tiredness (1–7)"
  ) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

irr
ti
ggsave("irr_daily_weeks.png", irr, width = 10, height = 6,
        dpi = 180)
ggsave("ti_daily_weeks.png", ti, width = 10, height = 6,
        dpi = 180)
```
# __ model 1 smooth overall
```{r}
model1 <- cmdstan_model(stan_file = root("stan", "gpbf1.stan"),
                        include_paths = root("stan"))
```
```{r}
standata1 <- list(x = daily$day,
                  y = log(daily$tired_mean),
                  N = length(daily$day),
                  c_f1 = 1.5,
                  M_f1 = 20)
```

```{r}
tic('Finding MAP for model 1 with optimization')
opt1 <- model1$optimize(data = standata1, init = 0, algorithm = "bfgs",
                        jacobian = TRUE)
```
```{r}
mytoc()
```

```{r}
odraws1 <- opt1$draws()
subset(odraws1, variable = c("intercept", "sigma_f1", "lengthscale_f1", "sigma")) |>
  as.data.frame() |>
  tt()
```

```{r}
oEf <- exp(as.numeric(subset(odraws1, variable = "f")))
make_pf(daily, oEf)
```
# __ model 2 with monthly periodic
```{r}
model2 <- cmdstan_model(stan_file = root("stan", "gpbf2.stan"),
                        include_paths = root("stan"))
```

```{r}
standata2 <- list(x = daily$day,
                  y = log(daily$tired_mean),
                  N = length(daily$day),
                  c_f1 = 1.5,
                  M_f1 = 20,
                  J_f2 = 20)
tic('Sampling from Pathfinder approximation of model 2 posterior')
pth2 <- model2$pathfinder(data = standata2, init = 0.1,
                          num_paths = 10, single_path_draws = 40, draws = 400,
                          history_size = 50, max_lbfgs_iters = 100,
                          refresh = 0)

mytoc()

pdraws2 <- pth2$draws()
summarise_draws(subset(pdraws2, variable = c("lp__")), n_distinct) |>
  tt()

summarise_draws(subset(pdraws2, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE),
                default_summary_measures()) |>
  tt()
```

```{r}
draws2 <- as_draws_matrix(pdraws2)
Ef <- exp(apply(subset(draws2, variable = "f"), 2, median))
Ef1 <- apply(subset(draws2, variable = "f1"), 2, median)
Ef1 <- exp(Ef1 - mean(Ef1) + mean(log(daily$tired_mean)))
Ef2 <- apply(subset(draws2, variable = "f2"), 2, median)
Ef2 <- exp(Ef2 - mean(Ef2) + mean(log(daily$tired_mean)))
pf  <- make_pf(daily, Ef) 
pf1 <- make_pf1(daily, Ef1)
pf2 <- make_pf2(daily, Ef2)
compose_3panel(pf, pf1, pf2)
```
# __ model 3 with day of week effect

```{r}
model3 <- cmdstan_model(stan_file = root("stan", "gpbf3.stan"),
                        include_paths = root("stan"))

standata3 <- list(x = daily$day,
                  y = log(daily$tired_mean),
                  N = length(daily$day),
                  c_f1 = 1.5, # factor c of basis functions for GP for f1
                  M_f1 = 20,  # number of basis functions for GP for f1
                  J_f2 = 20,  # number of basis functions for periodic f2
                  day_of_week = daily$wday_num)

tic('Sampling from Pathfinder approximation of model 3 posterior')
pth3 <- model3$pathfinder(data = standata3, init = 0.1,
                          num_paths = 10, single_path_draws = 40, draws = 400,
                          history_size = 50, max_lbfgs_iters = 100, 
                          refresh = 0)

pdraws3 <- pth3$draws()
summarise_draws(subset(pdraws3, variable = c("lp__")), n_distinct) |>
  tt()

summarise_draws(subset(pdraws3, variable = c("sigma_", "lengthscale_", "sigma", "beta_f3"), regex = TRUE),
                default_summary_measures()) |>
  tt()

draws3 <- as_draws_matrix(pdraws3)
Ef <- exp(apply(subset(draws3, variable = "f"), 2, median))
Ef1 <- apply(subset(draws3, variable = "f1"), 2, median)
Ef1 <- exp(Ef1 - mean(Ef1) + mean(log(daily$tired_mean)))
Ef2 <- apply(subset(draws3, variable = "f2"), 2, median)
Ef2 <- exp(Ef2 - mean(Ef2) + mean(log(daily$tired_mean)))
Ef_day_of_week <- apply(subset(draws3, variable = "f_day_of_week"), 2, median)
Ef_day_of_week <- exp(Ef_day_of_week - mean(Ef_day_of_week) + mean(log(daily$tired_mean)))
pf  <- make_pf(daily, Ef)
pf1 <- make_pf1(daily, Ef1)
pf2 <- make_pf2(daily, Ef2)
pf3 <- make_pf3(daily, Ef_day_of_week)
compose_4panel(pf, pf1, pf2, pf3)
```
# __ model 4 with varying weekday effect
```{r}
model4 <- cmdstan_model(stan_file = root("stan", "gpbf4.stan"),
                        include_paths = root("stan"))

standata4 <- list(x = daily$day,
                  y = log(daily$tired_mean),
                  N = length(daily$day),
                  c_f1 = 1.5, # factor c of basis functions for GP for f1
                  M_f1 = 20,  # number of basis functions for GP for f1
                  J_f2 = 20,  # number of basis functions for periodic f2
                  c_g3 = 1.5, # factor c of basis functions for GP for g3
                  M_g3 = 5,   # number of basis functions for GP for g3
                  day_of_week = daily$wday_num,
                  day_of_year = daily$day_of_year) 

tic('Sampling from Pathfinder approximation of model 4 posterior')
pth4 <- model4$pathfinder(data = standata4, init = 0.1,
                          num_paths = 10, single_path_draws = 40, draws = 400,
                          history_size = 50, max_lbfgs_iters = 100,
                          refresh = 0)

pdraws4 <- pth4$draws()
summarise_draws(subset(pdraws4, variable = c("lp__")), n_distinct) |>
  tt()

summarise_draws(subset(pdraws4, variable = c("sigma_", "lengthscale_", "sigma", "beta_f3"), regex = TRUE),
                default_summary_measures()) |>
  tt()

draws4 <- as_draws_matrix(pdraws4)
Ef <- exp(apply(subset(draws4, variable = "f"), 2, median))
Ef1 <- apply(subset(draws4, variable = "f1"), 2, median)
Ef1 <- exp(Ef1 - mean(Ef1) + mean(log(daily$tired_mean)))
Ef2 <- apply(subset(draws4, variable = "f2"), 2, median)
Ef2 <- exp(Ef2 - mean(Ef2) + mean(log(daily$tired_mean)))
Ef_day_of_week <- apply(subset(draws4, variable = "f_day_of_week"), 2, median)
Ef_day_of_week <- exp(Ef_day_of_week - mean(Ef_day_of_week) + mean(log(daily$tired_mean)))
Ef3 <- apply(subset(draws4, variable = "f3"), 2, median)
Ef3 <- exp(Ef3 - mean(Ef3) + mean(log(daily$tired_mean)))
pf   <- make_pf(daily, Ef)
pf1  <- make_pf1(daily, Ef1)
pf2  <- make_pf2(daily, Ef2)
pf3b <- make_pf3b(daily, Ef3)
compose_4panel(pf, pf1, pf2, pf3b)
```

```{r}
tic('MCMC sampling from model 4 posterior with Pathfinder initialization')
fit4 <- model4$sample(data = standata4, iter_warmup = 100, iter_sampling = 100,
                      chains = 4, parallel_chains = 4,
                      init = pth4)

mytoc()

draws4 <- fit4$draws()
summarise_draws(subset(draws4, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE)) |>
  tt()

summarise_draws(subset(draws4, variable = c("beta_f3"))) |>
  tt()

draws4 <- as_draws_matrix(draws4)
Ef <- exp(apply(subset(draws4, variable = "f"), 2, median))
Ef1 <- apply(subset(draws4, variable = "f1"), 2, median)
Ef1 <- exp(Ef1 - mean(Ef1) + mean(log(daily$tired_mean)))
Ef2 <- apply(subset(draws4, variable = "f2"), 2, median)
Ef2 <- exp(Ef2 - mean(Ef2) + mean(log(daily$tired_mean)))
Ef_day_of_week <- apply(subset(draws4, variable = "f_day_of_week"), 2, median)
Ef_day_of_week <- exp(Ef_day_of_week - mean(Ef_day_of_week) + mean(log(daily$tired_mean)))
Ef3 <- apply(subset(draws4, variable = "f3"), 2, median)
Ef3 <- exp(Ef3 - mean(Ef3) + mean(log(daily$tired_mean)))
pf   <- make_pf(daily, Ef) 
pf1  <- make_pf1(daily, Ef1)
pf2  <- make_pf2(daily, Ef2, date_breaks = "2 month")
pf3b <- make_pf3b(daily, Ef3 , weekday_labels = TRUE)
p<- compose_4panel(pf, pf1, pf2, pf3b)

ggplot(daily[i1,])+ geom_point(aes(day[i1], "y" = Ef3[i1], color = "red")) + geom_point(aes(x = day[i1+1], y = Ef3[i1+1], color = "orange")) + geom_point(aes(x = day[i1+2], y = Ef3[i1+2], color = "yellow")) + geom_point(aes(x = day[i1+3], y = Ef3[i1+3], colour = "green"))+ geom_point(aes(x = day[i1+4], y = Ef3[i1+4], colour = "blue"))+ geom_point(aes(x = day[i1+5], y = Ef3[i1+5], colour = "purple"))+ geom_point(aes(x = day[i1+6], y = Ef3[i1+6], colour = "black"))

ggsave("stan_fit_4.png", p, width = 10, height = 6,dpi = 180)
```

```{r}
daily$Ef3 <- round(Ef3, 2)

mondays <- daily[daily$wday == "Mon","Ef3"]
nrow(distinct(mondays))
tu <- daily[daily$wday == "Tue","Ef3"]
nrow(distinct(tu))
wed <- daily[daily$wday == "Wed","Ef3"]
nrow(distinct(wed))
thu <- daily[daily$wday == "Thu","Ef3"]
nrow(distinct(thu))
fri <- daily[daily$wday == "Fri","Ef3"]
nrow(distinct(fri))
sat <- daily[daily$wday == "Sat","Ef3"]
nrow(distinct(sat))
sun <- daily[daily$wday == "Sun","Ef3"]
nrow(distinct(sun))
plot(daily$wday, daily$Ef3)
plot(daily$wday, daily$tired_mean)
```
```
f_day_of_week[1] f_day_of_week[2] f_day_of_week[3] f_day_of_week[4] f_day_of_week[5] 
  mon   1.892175   tue 1.885922     wed 1.888777     thu 1.894157    fri  1.955837 
f_day_of_week[6] f_day_of_week[7] 
  sat   1.940643   sun   1.878867 
```
# model 3 is better, keeping it
```{r}
tic('MCMC sampling from model 3 posterior with Pathfinder initialization')
fit3 <- model3$sample(data = standata3, iter_warmup = 100, iter_sampling = 100,
                      chains = 4, parallel_chains = 4,
                      init = pth3)
```

```{r}
draws3 <- fit3$draws()
summarise_draws(subset(draws3, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE)) |>
  tt()

summarise_draws(subset(draws3, variable = c("beta_f3"))) |>
  tt()

draws3 <- as_draws_matrix(draws3)
Ef <- exp(apply(subset(draws3, variable = "f"), 2, median))
Ef1 <- apply(subset(draws3, variable = "f1"), 2, median)
Ef1 <- exp(Ef1 - mean(Ef1) + mean(log(daily$tired_mean)))
Ef2 <- apply(subset(draws3, variable = "f2"), 2, median)
Ef2 <- exp(Ef2 - mean(Ef2) + mean(log(daily$tired_mean)))
Ef_day_of_week <- apply(subset(draws3, variable = "f_day_of_week"), 2, median)
Ef_day_of_week <- exp(Ef_day_of_week - mean(Ef_day_of_week) + mean(log(daily$tired_mean)))
pf  <- make_pf(daily, Ef)
pf1 <- make_pf1(daily, Ef1)
pf2 <- make_pf2(daily, Ef2, date_breaks = "2 month")
pf3 <- make_pf3(daily, Ef_day_of_week)
compose_4panel(pf, pf1, pf2, pf3)
```


# comparison, sometime later
```{r}
loo1 <- fit1$loo()
loo2 <- fit2$loo()
loo3 <- fit3$loo()
loo4 <- fit4$loo()
loo6 <- fit6$loo()
loo7 <- fit7$loo()
loo_compare(list(`Model 1` = loo1, `Model 2` = loo2, `Model 3` = loo3, `Model 4` = loo4, `Model 6` = loo6, `Model 7` = loo7, `Model 8 + RHS` = loo8rhs)) |>
  as.data.frame() |>
  rownames_to_column("model") |>
  select(model, elpd_diff, se_diff) |>
  tt()


draws8 <- as_draws_matrix(draws8tnu)
f <- exp(subset(draws8, variable = "f"))
loo8tnu <- fit8tnu$loo(save_psis = TRUE)
Efloo <- E_loo(f, psis_object = loo8tnu$psis_object)$value
LOOR2 <- 1 - var(log(birthdays$births_relative100 / Efloo)) / var(log(birthdays$births_relative100))
data.frame(Model = "Model 8 Student's t", `LOO-R2` = LOOR2) |>
  tt()
```
# __ model 3_5 with day of week effect via a periodic

```{r}
model3_5 <- cmdstan_model(stan_file = root("stan", "gpbf3_5.stan"),include_paths = root("stan"))

standata3_5 <- list(x = daily$day,
                  y = log(daily$tired_mean),
                  N = length(daily$day),
                  c_f1 = 1.5, # factor c of basis functions for GP for f1
                  M_f1 = 20,  # number of basis functions for GP for f1
                  J_f2 = 20,  # number of basis functions for periodic f2
                  J_f3 = 20)

tic('Sampling from Pathfinder approximation of model 3 posterior')
pth3_5 <- model3_5$pathfinder(data = standata3_5, init = 0.1,
                          num_paths = 10, single_path_draws = 40, draws = 400,
                          history_size = 50, max_lbfgs_iters = 100, 
                          refresh = 0, psis_resample=FALSE)

pdraws3_5 <- pth3_5$draws()
summarise_draws(subset(pdraws3_5, variable = c("lp__")), n_distinct) |>
  tt()

summarise_draws(subset(pdraws3_5, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE),
                default_summary_measures()) |>
  tt()

draws3_5 <- as_draws_matrix(pdraws3_5)
Ef <- exp(apply(subset(draws3_5, variable = "f"), 2, median))
Ef1 <- apply(subset(draws3_5, variable = "f1"), 2, median)
Ef1 <- exp(Ef1 - mean(Ef1) + mean(log(daily$tired_mean)))
Ef2 <- apply(subset(draws3_5, variable = "f2"), 2, median)
Ef2 <- exp(Ef2 - mean(Ef2) + mean(log(daily$tired_mean)))
Ef3 <- apply(subset(draws3_5, variable = "f3"), 2, median)
Ef3 <- exp(Ef3 - mean(Ef3) + mean(log(daily$tired_mean)))
#Ef_day_of_week <- apply(subset(draws3_5, variable = "f_day_of_week"), 2, median)
#Ef_day_of_week <- exp(Ef_day_of_week - mean(Ef_day_of_week) + mean(log(daily$tired_mean)))
pf  <- make_pf(daily, Ef)
pf1 <- make_pf1(daily, Ef1)
pf2 <- make_pf2(daily, Ef2)
pf3 <- make_pf3(daily, Ef3)
compose_4panel(pf, pf1, pf2, pf3)
```

```{r}
tic('MCMC sampling from model 3_5 posterior with Pathfinder initialization')
fit3 <- model3_5$sample(data = standata3_5, iter_warmup = 100, iter_sampling = 100, chains = 4, parallel_chains = 4,  init = pth3_5)
draws3_5 <- fit3$draws()
summarise_draws(subset(draws3_5, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE)) |>
  tt()

draws3_5 <- as_draws_matrix(draws3_5)
Ef <- exp(apply(subset(draws3_5, variable = "f"), 2, median))
Ef1 <- apply(subset(draws3_5, variable = "f1"), 2, median)
Ef1 <- exp(Ef1 - mean(Ef1) + mean(log(daily$tired_mean)))
Ef2 <- apply(subset(draws3_5, variable = "f2"), 2, median)
Ef2 <- exp(Ef2 - mean(Ef2) + mean(log(daily$tired_mean)))
Ef3 <- apply(subset(draws3_5, variable = "f3"), 2, median)
Ef3 <- exp(Ef3 - mean(Ef3) + mean(log(daily$tired_mean)))
pf  <- make_pf(daily, Ef)
pf1 <- make_pf1(daily, Ef1)
pf2 <- make_pf2(daily, Ef2, date_breaks = "2 month")
pf3 <- make_pf3(daily, Ef3)
compose_4panel(pf, pf1, pf2, pf3)
```

```{r}
ggsave("stan_final.png", compose_4panel(pf, pf1, pf2, pf3), width = 10, height = 6,
        dpi = 180)
save(fit3, file = "fit3_5.RData")
```
# model 3_5 with irritation instead
```{r}
model3_5 <- cmdstan_model(stan_file = root("stan", "gpbf3_5.stan"),include_paths = root("stan"))

standata3_5 <- list(x = daily$day,
                  y = log(daily$irr_mean),
                  N = length(daily$day),
                  c_f1 = 1.5, # factor c of basis functions for GP for f1
                  M_f1 = 20,  # number of basis functions for GP for f1
                  J_f2 = 20,  # number of basis functions for periodic f2
                  J_f3 = 20)

tic('Sampling from Pathfinder approximation of model 3 posterior')
pth3_5 <- model3_5$pathfinder(data = standata3_5, init = 0.1,
                          num_paths = 10, single_path_draws = 40, draws = 400,
                          history_size = 50, max_lbfgs_iters = 100, 
                          refresh = 0, psis_resample=FALSE)

pdraws3_5 <- pth3_5$draws()
summarise_draws(subset(pdraws3_5, variable = c("lp__")), n_distinct) |>
  tt()

summarise_draws(subset(pdraws3_5, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE),
                default_summary_measures()) |>
  tt()

draws3_5 <- as_draws_matrix(pdraws3_5)
Ef <- exp(apply(subset(draws3_5, variable = "f"), 2, median))
Ef1 <- apply(subset(draws3_5, variable = "f1"), 2, median)
Ef1 <- exp(Ef1 - mean(Ef1) + mean(log(daily$irr_mean)))
Ef2 <- apply(subset(draws3_5, variable = "f2"), 2, median)
Ef2 <- exp(Ef2 - mean(Ef2) + mean(log(daily$irr_mean)))
Ef3 <- apply(subset(draws3_5, variable = "f3"), 2, median)
Ef3 <- exp(Ef3 - mean(Ef3) + mean(log(daily$irr_mean)))
#Ef_day_of_week <- apply(subset(draws3_5, variable = "f_day_of_week"), 2, median)
#Ef_day_of_week <- exp(Ef_day_of_week - mean(Ef_day_of_week) + mean(log(daily$irr_mean)))
pf  <- make_pf(daily, Ef)
pf1 <- make_pf1(daily, Ef1)
pf2 <- make_pf2(daily, Ef2)
pf3 <- make_pf3(daily, Ef3)
compose_4panel(pf, pf1, pf2, pf3)
```
```{r}
tic('MCMC sampling from model 3_5 posterior with Pathfinder initialization')
fit3_5 <- model3_5$sample(data = standata3_5, iter_warmup = 100, iter_sampling = 100, chains = 4, parallel_chains = 4,  init = pth3_5)
draws3_5 <- fit3_5$draws()
summarise_draws(subset(draws3_5, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE)) |>
  tt()

draws3_5 <- as_draws_matrix(draws3_5)
Ef <- exp(apply(subset(draws3_5, variable = "f"), 2, median))
Ef1 <- apply(subset(draws3_5, variable = "f1"), 2, median)
Ef1 <- exp(Ef1 - mean(Ef1) + mean(log(daily$irr_mean)))
Ef2 <- apply(subset(draws3_5, variable = "f2"), 2, median)
Ef2 <- exp(Ef2 - mean(Ef2) + mean(log(daily$irr_mean)))
Ef3 <- apply(subset(draws3_5, variable = "f3"), 2, median)
Ef3 <- exp(Ef3 - mean(Ef3) + mean(log(daily$irr_mean)))
pf  <- make_pf(daily, Ef)
pf1 <- make_pf1(daily, Ef1)
pf2 <- make_pf2(daily, Ef2, date_breaks = "2 month")
pf3 <- make_pf3(daily, Ef3)
compose_4panel(pf, pf1, pf2, pf3)
```
# different priors
```{r}
model3_5_5 <- cmdstan_model(stan_file = root("stan", "gpbf3_5_5.stan"),include_paths = root("stan"))

standata3_5_5 <- list(x = daily$day,
                  y = log(daily$irr_mean),
                  N = length(daily$day),
                  c_f1 = 1.5, # factor c of basis functions for GP for f1
                  M_f1 = 20,  # number of basis functions for GP for f1
                  J_f2 = 20,  # number of basis functions for periodic f2
                  J_f3 = 20)

tic('Sampling from Pathfinder approximation of model 3 posterior')
pth3_5_5 <- model3_5_5$pathfinder(data = standata3_5_5, init = 0.1,
                          num_paths = 10, single_path_draws = 40, draws = 400,
                          history_size = 50, max_lbfgs_iters = 100, 
                          refresh = 0, psis_resample=FALSE)

pdraws3_5_5 <- pth3_5_5$draws()
summarise_draws(subset(pdraws3_5_5, variable = c("lp__")), n_distinct) |>
  tt()

summarise_draws(subset(pdraws3_5_5, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE),
                default_summary_measures()) |>
  tt()

draws3_5_5 <- as_draws_matrix(pdraws3_5_5)
Ef <- exp(apply(subset(draws3_5_5, variable = "f"), 2, median))
Ef1 <- apply(subset(draws3_5_5, variable = "f1"), 2, median)
Ef1 <- exp(Ef1 - mean(Ef1) + mean(log(daily$irr_mean)))
Ef2 <- apply(subset(draws3_5_5, variable = "f2"), 2, median)
Ef2 <- exp(Ef2 - mean(Ef2) + mean(log(daily$irr_mean)))
Ef3 <- apply(subset(draws3_5_5, variable = "f3"), 2, median)
Ef3 <- exp(Ef3 - mean(Ef3) + mean(log(daily$irr_mean)))
#Ef_day_of_week <- apply(subset(draws3_5, variable = "f_day_of_week"), 2, median)
#Ef_day_of_week <- exp(Ef_day_of_week - mean(Ef_day_of_week) + mean(log(daily$irr_mean)))
pf  <- make_pf(daily, Ef)
pf1 <- make_pf1(daily, Ef1)
pf2 <- make_pf2(daily, Ef2)
pf3 <- make_pf3(daily, Ef3)
compose_4panel(pf, pf1, pf2, pf3)
```
# todo detrending

```{r}
df

mean_irr <- mean(df$irritated)
mean_tir <- mean(df$tired)

df$irritated <- df$irritated - mean_irr
df$tired <- df$tired - mean_tir

daily_d <- df %>%
  group_by(date_parsed, year, month, month_num, week, wday, wday_num) %>%
  summarise(
    tired_mean = mean(tired),
    tired_sd   = sd(tired),
    irr_mean = mean(irritated),
    irr_sd = sd(irritated),
    n_beeps   = n(),
    .groups   = "drop"
  )

daily_d <- daily_d %>% mutate(
  timestamp = as.numeric(date_parsed),
  day = row_number(),
  day_of_year = as.numeric(strftime(date_parsed, format = "%j")),
  day_of_month = day(date_parsed)
)

p_irr <- daily_d |>  ggplot(aes(x = day, y = irr_mean)) +geom_point() + hline_0(color = "red")
p_tir <- daily_d |>  ggplot(aes(x = day, y = tired_mean)) +geom_point()+ hline_0(color = "red")
p_irr
p_tir
```
# detrended values, removed logs and exps, pathfinder is the same
```{r}
model3_5 <- cmdstan_model(stan_file = root("stan", "gpbf3_5.stan"),include_paths = root("stan"))

standata3_5 <- list(x = daily_d$day,
                  y = (daily_d$irr_mean),
                  N = length(daily_d$day),
                  c_f1 = 1.5, # factor c of basis functions for GP for f1
                  M_f1 = 20,  # number of basis functions for GP for f1
                  J_f2 = 20,  # number of basis functions for periodic f2
                  J_f3 = 20)

tic('Sampling from Pathfinder approximation of model 3 posterior')
pth3_5 <- model3_5$pathfinder(data = standata3_5, init = 0.1,
                          num_paths = 10, single_path_draws = 40, draws = 400,
                          history_size = 50, max_lbfgs_iters = 100, 
                          refresh = 0, psis_resample=FALSE)

pdraws3_5 <- pth3_5$draws()
summarise_draws(subset(pdraws3_5, variable = c("lp__")), n_distinct) |>
  tt()

summarise_draws(subset(pdraws3_5, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE),
                default_summary_measures()) |>
  tt()

draws3_5 <- as_draws_matrix(pdraws3_5)
Ef <- (apply(subset(draws3_5, variable = "f"), 2, median))
Ef1 <- apply(subset(draws3_5, variable = "f1"), 2, median)
Ef1 <- (Ef1 - mean(Ef1) + mean((daily_d$irr_mean)))
Ef2 <- apply(subset(draws3_5, variable = "f2"), 2, median)
Ef2 <- (Ef2 - mean(Ef2) + mean((daily_d$irr_mean)))
Ef3 <- apply(subset(draws3_5, variable = "f3"), 2, median)
Ef3 <- (Ef3 - mean(Ef3) + mean((daily_d$irr_mean)))
#Ef_day_of_week <- apply(subset(draws3_5, variable = "f_day_of_week"), 2, median)
#Ef_day_of_week <- exp(Ef_day_of_week - mean(Ef_day_of_week) + mean(log(daily_d$irr_mean)))
pf  <- make_pf(daily_d, Ef)
pf1 <- make_pf1(daily_d, Ef1)
pf2 <- make_pf2(daily_d, Ef2)
pf3 <- make_pf3(daily_d, Ef3)
compose_4panel(pf, pf1, pf2, pf3)

```

```{r}
tic('MCMC sampling from model 3_5 posterior with Pathfinder initialization')
fit3_5 <- model3_5$sample(data = standata3_5, iter_warmup = 100, iter_sampling = 100, chains = 4, parallel_chains = 4,  init = pth3_5)
draws3_5 <- fit3_5$draws()
summarise_draws(subset(draws3_5, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE)) |>
  tt()

draws3_5 <- as_draws_matrix(draws3_5)
Ef <- (apply(subset(draws3_5, variable = "f"), 2, median))
Ef1 <- apply(subset(draws3_5, variable = "f1"), 2, median)
#Ef1 <- (Ef1 - mean(Ef1) + mean((daily_d$irr_mean)))
Ef2 <- apply(subset(draws3_5, variable = "f2"), 2, median)
#Ef2 <- (Ef2 - mean(Ef2) + mean((daily_d$irr_mean)))
Ef3 <- apply(subset(draws3_5, variable = "f3"), 2, median)
#Ef3 <- (Ef3 - mean(Ef3) + mean((daily_d$irr_mean)))
pf  <- make_pf(daily_d, Ef)
pf1 <- make_pf1(daily_d, Ef1)
pf2 <- make_pf2(daily_d, Ef2, date_breaks = "2 month")
pf3 <- make_pf3(daily_d, Ef3)
compose_4panel(pf, pf1, pf2, pf3)
```

```{r}
ggsave("stan_final_detrended.png", compose_4panel(pf, pf1, pf2, pf3), width = 10, height = 6,
        dpi = 180)
save(fit3_5, file = "fit3_5_irr_detrended.RData")
load(file = "fit3_5_irr_detrended.RData")
```

# no pathfinder init
```{r}
#tic('MCMC sampling from model 3_5 posterior with Pathfinder initialization')
fit3_5_n <- model3_5$sample(data = standata3_5, iter_warmup = 50, iter_sampling = 50, chains = 3, parallel_chains = 3)#,  init = pth3_5)
draws3_5_n <- fit3_5_n$draws()
summarise_draws(subset(draws3_5_n, variable = c("sigma_", "lengthscale_", "sigma"), regex = TRUE)) |>
  tt()

draws3_5_n <- as_draws_matrix(draws3_5_n)
Ef <- (apply(subset(draws3_5_n, variable = "f"), 2, median))
Ef1 <- apply(subset(draws3_5_n, variable = "f1"), 2, median)
#Ef1 <- (Ef1 - mean(Ef1) + mean((daily_d$irr_mean)))
Ef2 <- apply(subset(draws3_5_n, variable = "f2"), 2, median)
#Ef2 <- (Ef2 - mean(Ef2) + mean((daily_d$irr_mean)))
Ef3 <- apply(subset(draws3_5_n, variable = "f3"), 2, median)
#Ef3 <- (Ef3 - mean(Ef3) + mean((daily_d$irr_mean)))
pf  <- make_pf(daily_d, Ef)
pf1 <- make_pf1(daily_d, Ef1)
pf2 <- make_pf2(daily_d, Ef2, date_breaks = "2 month")
pf3 <- make_pf3(daily_d, Ef3)
compose_4panel(pf, pf1, pf2, pf3)
```
# simple model
```{r}
library(dplyr)
library(rstan)
library(ggplot2)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())


stan_data <- list(
  N = nrow(daily_d),
  y = daily_d$irr_mean,
  x = daily_d$day
)

fit_eq_overall <- stan(
  file = "C:/Users/zubareva/OneDrive - Tilburg University/Desktop/code/phd/data examples/stan/gp_eq_overall.stan",
  data = stan_data,
  chains = 4,
  iter = 500,
  warmup = 200,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_eq_overall,
  pars = c("length_scale","variance","sigma","eta")
)

save(
  fit_eq_overall,
  file = "fit_eq_overall.RData"
)


post_eq_overall <- rstan::extract(fit_eq_overall)
eta <- apply(post_eq_overall$eta, 2, mean)
plot(daily_d$day, eta)
```
```{r}

stan_data <- list(
  N1 = nrow(daily_d),
  y1 = daily_d$irr_mean,
  x1 = daily_d$day,
  N2 = 62,
  x2 = seq(239, 300, by = 1)
)

fit_eq_overall_pr <- stan(
  file = "C:/Users/zubareva/OneDrive - Tilburg University/Desktop/code/phd/data examples/stan/gp_eq_overall_predictive.stan",
  data = stan_data,
  chains = 4,
  iter = 500,
  warmup = 200,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_eq_overall_pr,
  pars = c("length_scale","variance","sigma","eta")
)

# save(
#   fit_eq_overall_pr,
#   file = "fit_eq_overall_pr.RData"
# )


post_eq_overall_pr <- rstan::extract(fit_eq_overall_pr)
y2 <- apply(post_eq_overall_pr$y2, 2, mean)
pred <- data.frame(x = append(daily_d$day, seq(239, 300, by = 1)), y = append(daily_d$irr_mean, y2))
plot( seq(239, 300, by = 1), y2)
plot(pred$x, pred$y)
```

# check fit
```{r}
stan_data <- list(
  N = nrow(daily_d),
  y = daily_d$irr_mean,
  x = daily_d$day
)

fit_eq_overall <- stan(
  file = "C:/Users/zubareva/OneDrive - Tilburg University/Desktop/code/phd/data examples/stan/gp_eq_overall.stan",
  data = stan_data,
  chains = 4,
  iter = 500,
  warmup = 200,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_eq_overall,
  pars = c("length_scale","variance","sigma","eta", "fit")
)

# save(
#   fit_eq_overall,
#   file = "fit_eq_overall.RData"
# )


post_eq_overall <- rstan::extract(fit_eq_overall)
fit <- apply(post_eq_overall$fit, 2, mean)
par(mfrow = c(1, 2))
p1<-plot(daily_d$day, fit)
p2<-plot(daily_d$day, daily_d$irr_mean)
p1
p2
```

```{r}
var( daily_d$irr_mean)
par(mfrow = c(1, 2))
p1<-plot(daily_d$day, fit, type = "l")
p2<-plot(daily_d$day, (daily_d$irr_mean)/(var(daily_d$irr_mean)))
p1
p2
```
# check fit of synthetic data
```{r}
x<- seq(1, 100, by=1)
y<- 30*sin(x/5) #+ x
plot(x, y, type = "l")

stan_data <- list(
  N = length(x),
  y = y,
  x = x
)

fit_eq_overall <- stan(
  file = "C:/Users/zubareva/OneDrive - Tilburg University/Desktop/code/phd/data examples/stan/gp_eq_overall.stan",
  data = stan_data,
  chains = 4,
  iter = 500,
  warmup = 200,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_eq_overall,
  pars = c("length_scale","variance","sigma","eta", "fit")
)

# save(
#   fit_eq_overall,
#   file = "fit_eq_overall.RData"
# )

post_eq_overall$length_scale

post_eq_overall <- rstan::extract(fit_eq_overall)
fit <- apply(post_eq_overall$fit, 2, mean)
eta <- apply(post_eq_overall$eta, 2, mean)
#par(mfrow = c(1, 2))
p1<-plot(x, fit)
plot(x, eta)
#p2<-plot(x, y)
p1
#p2
```
# fit is ok even for large scales, will try different priors

```{r}
stan_data <- list(
  N = length(x),
  y = y,
  x = x
)

fit_eq_overall <- stan(
  file = "C:/Users/zubareva/OneDrive - Tilburg University/Desktop/code/phd/data examples/stan/gp_eq_overall_weird_priors.stan",
  data = stan_data,
  chains = 4,
  iter = 500,
  warmup = 200,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_eq_overall,
  pars = c("length_scale","variance","sigma","eta", "fit")
)

# save(
#   fit_eq_overall,
#   file = "fit_eq_overall.RData"
# )


post_eq_overall <- rstan::extract(fit_eq_overall)
fit <- apply(post_eq_overall$fit, 2, mean)
eta <- apply(post_eq_overall$eta, 2, mean)
pred<- data.frame(x = x, y = y, fit = fit)
ggplot(pred, aes(x, y)) + geom_point() + geom_line(aes(x, fit), color = "red")
```
# weid priors are not breaking synthetic data

# weird priors + prediction work 
#  intercept ~ normal(0, 5);
#  length_scale ~ inv_gamma(5, 1);
#  variance ~ normal(100, 50);
#  sigma ~ normal(1000, 50);
#  eta ~ normal(100, 10);

```{r}
stan_data <- list(
  N1 = nrow(daily_d),
  y1 = daily_d$irr_mean,
  x1 = daily_d$day,
  N2 = 62,
  x2 = seq(239, 300, by = 1)
)

fit_eq_overall_pr <- stan(
  file = "C:/Users/zubareva/OneDrive - Tilburg University/Desktop/code/phd/data examples/stan/gp_eq_overall_predictive.stan",
  data = stan_data,
  chains = 4,
  iter = 500,
  warmup = 200,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_eq_overall_pr,
  pars = c("length_scale","variance","sigma", "intercept")
)

save(
  fit_eq_overall_pr,
  file = "what.RData"
)

load(file = "what.RData")

new_fit <- rstan::extract(fit_eq_overall_pr)
fit_n <- apply(new_fit$fit, 2, mean)
y2_n <- apply(new_fit$y2, 2, mean)
pred <- data.frame(da_x = append(daily_d$day, seq(239, 300, by = 1)), 
                   da_y = append(daily_d$irr_mean, rep(0, 62)), 
                   gp = append(fit_n, y2_n))

ggplot(pred, aes(da_x, da_y)) + geom_point() + geom_line(aes(da_x, gp), color = "red")
beep()
```
# not-squared params
```{r}

stan_data <- list(
  N = nrow(daily_d),
  y = daily_d$irr_mean,
  x = daily_d$day
)

fit_nonsq <- stan(
  file = "C:/Users/zubareva/OneDrive - Tilburg University/Desktop/code/phd/data examples/stan/gp_eq_overall.stan",
  data = stan_data,
  chains = 3,
  iter = 300,
  warmup = 100,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_nonsq,
  pars = c("length_scale","variance","sigma","intercept")
)

save(
  fit_nonsq,
  file = "fit_nonsq.RData"
)


post_fit_nonsq <- rstan::extract(fit_nonsq)

fit_n <- apply(post_fit_nonsq$fit, 2, mean)

pred <- data.frame(da_x = (daily_d$day), 
                   da_y = (daily_d$irr_mean), 
                   gp = (fit_n))

p<-ggplot(pred, aes(da_x, da_y)) + geom_point() + geom_line(aes(da_x, gp), color = "red")
p
ggsave("fiiiinal.png", p, width = 10, height = 6,
        dpi = 180)
```

# combined eq and  periodic

```{r}
stan_data <- list(
  N = nrow(daily_d),
  y = daily_d$irr_mean,
  x = daily_d$day
)

fit_eq_per <- stan(
  file = "C:/Users/margo/stan/gp_eq_periodic_weekly.stan",
  data = stan_data,
  chains = 4,
  iter = 500,
  warmup = 200,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_eq_per,
  pars = c("length_scale_k1","variance_k1", "length_scale_k2","variance_k2","sigma","intercept")
)

save(
  fit_eq_per,
  file = "fit_eq_per.RData"
)


post_fit_eq_per <- rstan::extract(fit_eq_per)

fit_n <- apply(post_fit_eq_per$fit, 2, mean)

pred <- data.frame(da_x = (daily_d$day), 
                   da_y = (daily_d$irr_mean), 
                   gp = (fit_n))

ggplot(pred, aes(da_x, da_y)) + geom_point() + geom_line(aes(da_x, gp), color = "red")
beep()
```

```{r}
stan_data <- list(
  N = nrow(daily_d),
  y = daily_d$irr_mean,
  x = daily_d$day
)

fit_ou_per <- stan(
  file = "C:/Users/margo/stan/gp_OU_periodic_weekly.stan",
  data = stan_data,
  chains = 4,
  iter = 500,
  warmup = 200,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_ou_per,
  pars = c("length_scale_k1","variance_k1", "length_scale_k2","variance_k2","sigma","intercept")
)

save(
  fit_ou_per,
  file = "fit_ou_per.RData"
)


post_fit_ou_per <- rstan::extract(fit_ou_per)

fit_n <- apply(post_fit_ou_per$fit, 2, mean)

pred <- data.frame(da_x = (daily_d$day), 
                   da_y = (daily_d$irr_mean), 
                   gp = (fit_n))

ggplot(pred, aes(da_x, da_y)) + geom_point() + geom_line(aes(da_x, gp), color = "red")
beep()
```

```{r}
stan_data <- list(
  N = nrow(daily_d),
  y = daily_d$irr_mean,
  x = daily_d$day
)

fit_ou_per_per <- stan(
  file = "C:/Users/margo/stan/gp_OU_periodic_weekly_monthly.stan",
  data = stan_data,
  chains = 4,
  iter = 500,
  warmup = 200,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_ou_per_per,
  pars = c("length_scale_k1","variance_k1", "length_scale_k2","variance_k2","length_scale_k3","variance_k3","sigma","intercept")
)

save(
  fit_ou_per_per,
  file = "fit_ou_per_per.RData"
)


post_fit_ou_per_per <- rstan::extract(fit_ou_per_per)

fit_n <- apply(post_fit_ou_per_per$fit, 2, mean)

pred <- data.frame(da_x = (daily_d$day), 
                   da_y = (daily_d$irr_mean), 
                   gp = (fit_n))

ggplot(pred, aes(da_x, da_y)) + geom_point() + geom_line(aes(da_x, gp), color = "red")
beep()
```
# decomposed

```{r}

stan_data <- list(
  N = nrow(daily_d),
  y = daily_d$irr_mean,
  x = daily_d$day
)

fit_ou_per_per_dec <- stan(
  file = "C:/Users/margo/stan/gp_OU_periodic_weekly_monthly_decomposed.stan",
  data = stan_data,
  chains = 4,
  iter = 500,
  warmup = 200,
  seed = 123,
  control = list(adapt_delta = 0.9)
)

print(
  fit_ou_per_per_dec,
  pars = c("length_scale_k1","variance_k1", "length_scale_k2","variance_k2","length_scale_k3","variance_k3","sigma","intercept")
)

save(
  fit_ou_per_per_dec,
  file = "fit_ou_per_per_dec.RData"
)


post_fit_ou_per_per_dec <- rstan::extract(fit_ou_per_per_dec)

fit_n <- apply(post_fit_ou_per_per_dec$fit, 2, mean)
eff_1 <- apply(post_fit_ou_per_per_dec$eff_1, 2, mean)
eff_2 <- apply(post_fit_ou_per_per_dec$eff_2, 2, mean)
eff_3 <- apply(post_fit_ou_per_per_dec$eff_3, 2, mean)

pred <- data.frame(da_x = (daily_d$day), 
                   da_y = (daily_d$irr_mean), 
                   gp = (fit_n), 
                   e_1 = eff_1, 
                   e_2 = eff_2, 
                   e_3 = eff_3)

fit_plot<- ggplot(pred, aes(da_x, da_y)) + geom_point() + geom_line(aes(da_x, gp), color = "red")
eff_1_plot <- ggplot(pred, aes(da_x, da_y)) + geom_point() + geom_line(aes(da_x, e_1), color = "red")
eff_2_plot <- ggplot(pred, aes(da_x, da_y)) + geom_point() + geom_line(aes(da_x, e_2), color = "red")
eff_3_plot <- ggplot(pred, aes(da_x, da_y)) + geom_point() + geom_line(aes(da_x, e_3), color = "red")
(fit_plot + eff_1_plot) / (eff_2_plot + eff_3_plot)
beep()
```

