---
title: "gp_analysis"
output: html_document
date: "2026-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
########## Analysis for simulating GP with an artificial example
```{r}
#install.packages("tidyverse")
#install.packages("lgpr")
#install.packages("brms")
#install.packages("rstan")
#install.packages("rstanarm")
#install.packages("remotes")
#install.packages("tidybayes")
#install.packages("bridgesampling")
#install.packages("shinystan")
#install.packages("mgcv")
#remotes::install_github("michael-franke/aida-package")
#remotes::install_github("michael-franke/faintr")
#remotes::install_github("CogSciPrag/cspplot")

library(tidyverse)
library(dplyr)
library(lgpr)
library(brms)
library(rstan)
library(rstanarm)
library(remotes)
library(tidybayes)
library(bridgesampling)
library(shinystan)
library(mgcv)
library(aida)
library(faintr)
library(cspplot)

# these options help Stan run faster
options(mc.cores = parallel::detectCores())

# use the CSP-theme for plotting
theme_set(theme_csp())

# global color scheme from CSP
project_colors = cspplot::list_colors() |> pull(hex)
# names(project_colors) <- cspplot::list_colors() |> pull(name)

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
}
```

```{r}
get_GP_simulation <- 
  function(x = seq(0,10, by = 0.1), 
           Intercept = 0, 
           slope = 1, 
           sigma = 1, 
           sigma_f=0.5, 
           lambda=100, 
           seed = NULL) {
    
    if (! is.null(seed)){
      set.seed(seed)
    }
    
    # number of points to generate prediction for
    N <- length(x)
    
    # linear predictor (vanilla LM)
    eta = Intercept + slope * x
    
    # kernel function (here: radial basis function kernel)
    get_covmatrix <- function(x, sigma_f, lambda) {
      K = matrix(0, nrow=N, ncol=N)
      for (i in 1:N) {
        for (j in 1:N) {
          K[i,j] = sigma_f^2 * exp(sqrt((i-j)^2) / (-2 *lambda))
        }
      }
      return(K)
    }
    
    # covariance matrix
    K <- get_covmatrix(x, sigma_f, lambda)
    
    # Gaussian process wiggles
    epsilon_gp <- mvtnorm::rmvnorm(
      n     = 1, 
      mean  = rep(0, N),
      sigma = K)[1,]
    
    # central tendency
    mu <- epsilon_gp + eta
    
    # data prediction
    y <- rnorm(N, mu, sigma)
      
    tibble(x, eta, mu, y) 
  }

plot_GP_simulation <- function(GP_simulation) {
  GP_simulation |> 
    ggplot(aes(x = x, y = eta )) + 
    geom_line(color = project_colors[1], size = 1.25) +
    geom_line(aes(y = mu), color = project_colors[2], size = 1.25) +
    geom_point(aes(y = y), color = project_colors[3], alpha = 0.7, size = 1.2)
}
```

```{r}
get_GP_simulation(
  x = seq(-1,1, length.out = 500), 
  Intercept = 0, 
  slope = 0.3, 
  sigma_f = 0.05, 
  lambda  = 20, 
  sigma = 0.05) |> 
  plot_GP_simulation()
```

```{r}
GP_simulation <- get_GP_simulation(
  x = seq(-1,1, length.out = 100), 
  Intercept = 0, 
  slope = 1, 
  sigma_f = 0.5, 
  lambda  = 2, 
  sigma = 0.0001,
  seed = 1996)

plot_GP_simulation(GP_simulation)
```

```{r}
# results: hide
fit_GPR <- 
  brms::brm(
    formula = y ~ gp(x) + x,
    data    = GP_simulation,
    iter    = 2000
  )
```

```{r}
summary(fit_GPR)
```

```{r}
conditional_effects(fit_GPR)
```

########### Artificial data fits fine
############ Data of WHR for Benevolence in Netherlands
```{r}
x <- c(2007, 2008, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024)
y <- c(74.9, 77.1, 75.3, 74.0, 69.2, 70.8, 73.5, 66.6, 64.5, 65.8, 57.1, 62.3, 55.8, 68.1, 64.1, 64.0, 60.7)

data<- data.frame(x, y)

ggplot(data, aes(x = x, y = y )) + 
    geom_line(color = project_colors[1], size = 1.25) +
    geom_point(aes(y = y), color = project_colors[3], alpha = 0.7, size = 1.2)
```

```{r}
data$x_s <- (scale(data$x))[,1]

fit_GPR_2 <- brms::brm(
  y ~ gp(x_s, kappa = 1e-8),
  data = data,
  iter = 4000
)


summary(fit_GPR_2)
```

########### 17 observations are not enough

```{r}
newdata <- data.frame(
  x_s = seq(min(data$x_s), max(data$x_s), length.out = 200)
)

pred <- fitted(
  fit_GPR_2,
  newdata = newdata,
  re_formula = NA
)

```

############################################### Analysis of Japan panel on covid, 30 waves

```{r}
panel_japan <- read.csv("data.csv", stringsAsFactors = FALSE)

j_data<- panel_japan[, c("cw_ID", "cv_r1", "wave", "sleep_x", "only_jpn")]

j_data <- j_data %>%
  mutate(
    wave_s   = as.numeric(scale(wave)),
    #id       = as.integer(factor(cw_ID)),
    id       = as.factor(factor(cw_ID)),
    cv_r1    = as.numeric(scale(cv_r1)),
    sleep_x  = as.numeric(scale(sleep_x)),
    only_jpn = as.numeric(scale(as.numeric(only_jpn))
  ))

j_data<- j_data[, c("id", "cv_r1", "wave_s")]

head(j_data)
```

# BRMS analysis on just one person - 30 observations, does not work
```{r}
j_data_1 <- subset(j_data, id == "8491")

ggplot(j_data_1, aes(x = wave_s, y = cv_r1 )) + 
    geom_line(color = project_colors[1], size = 1.25) +
  geom_smooth(color = project_colors[2], size = 1) +  
  geom_point(aes(y = cv_r1), color = project_colors[3], alpha = 0.7, size = 1.2)

ggplot(j_data_1, aes(x = wave_s, y = sleep_x )) + 
    geom_line(color = project_colors[1], size = 1.25) +
    geom_smooth(color = project_colors[2], size = 1) + 
    geom_point(aes(y = sleep_x), color = project_colors[3], alpha = 0.7, size = 1.2)

ggplot(j_data_1, aes(x = wave_s, y = only_jpn )) + 
    geom_smooth(color = project_colors[2], size = 1) +
    geom_line(color = project_colors[1], size = 1.25) +
  geom_point(aes(y = only_jpn), color = project_colors[3], alpha = 0.7, size = 1.2)
```
##### just one person, 17 obs - not working
```{r}
fit_GPR_3 <- brms::brm(
  cv_r1 ~ gp(wave_s, kappa = 1e-2),
  data = j_data_1,
  iter = 4000
)

summary(fit_GPR_3)

newdata <- data.frame(
  wave_s = seq(min(j_data_1$wave_s), max(j_data_1$wave_s), length.out = 200)
)

pred <- fitted(
  fit_GPR_3,
  newdata = newdata,
  re_formula = NA
)
```
############################### Analysis with lgpr to study several people
```{r}
# For reproducibility
set.seed(123)

ggplot() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(x = "wave", y = "Dread") +
  theme(legend.position = "none") +
  theme(aspect.ratio=0.9) +
geom_smooth(data = j_data, aes(x=wave_s, y=cv_r1)) +
  ggtitle("Average dreadfulness levels") +
  theme(plot.title = element_text(size = 13, face = "bold")) 
```

```{r}
# For reproducibility
set.seed(123)

# Define mean and SD of the log-normal distribution
mean_priordistribution_lengthscale <- 0
sd_priordistribution_lengthscale <- 1

# Define warping function
warpfun <- function(time,warpingparameter){
  2 / (1 + exp( - warpingparameter * time)) - 1
}

# Define kernel with warping function
kernelGP <- function(X1, X2, lengthscale, alpha, warpingparameter, fun=function(x,w){x}){
  
  X1mat <- matrix(rep(fun(X1,warpingparameter),length(X2)),nrow=length(X1))
  X2mat <- matrix(rep(fun(X2,warpingparameter),length(X1)),nrow=length(X1),byrow=TRUE)
  
  alpha**2 * exp(- .5 * (X1mat - X2mat) * (X1mat - X2mat) / lengthscale**2 )
  
}

n <- 100 
X <- seq(-100, 100, length=n) 
mu = rep(0,length(X))

# Set up a 4x4 plotting panel
par(mfrow = c(4, 4), mar = c(2, 2, 2, 2))

# Loop to generate 16 plots
for (i in 1:16) {

# Lengthscale parameter
lengthscale1 <- dlnorm(1, meanlog = mean_priordistribution_lengthscale, sdlog = sd_priordistribution_lengthscale)

# Magnitude parameter (potentially modify prior distribution here!)
alpha1 <- dlnorm(1, meanlog = 0, sdlog = 1)

# Warping parameter (potentially modify prior distribution here!)
warpingparameter <- dlnorm(1, meanlog=-0.7,sdlog=0.9)
covm = kernelGP(X, X, lengthscale = lengthscale1, alpha = alpha1, warpingparameter = warpingparameter, fun = warpfun)

# Intercept
beta0 <- 0

  # Draw a line from the GP
  draw_f <- mvtnorm::rmvnorm(1, mean = mu, sigma = covm)
  
  # Plot the line
  plot(X, beta0 + draw_f, type = "l", col = "grey", lwd = 2,
       ylim = c(beta0 - 2 * alpha1, beta0 + 2 * alpha1),
       ylab = "y", xlab = "x", main = paste("Plot", i))
}

prior <- list(
  alpha = log_normal(0, 1), 
  ell = log_normal(0, 1),
  wrp = log_normal(-0.7, 0.9)
)
```


```{r}
j_sub <- subset(j_data, id %in% c("42040", "8491", "15569", "94338"))

#gp <- lgp(
#  cv_r1 ~
#    gp_ns(wave_s) +  # population-level GP
#    gp_ns(wave_s)*zs(id),# group-level GP
#  data = j_sub,
#  sample_f = TRUE
#)

gp_2 <- lgp(
  cv_r1 ~
    gp_ns(wave_s) +  zs(id),
  data = j_sub,
  sample_f = TRUE
)

summary(gp_2)

print(gp_2)
save(gp_2, file="gp_2.RData")
load("gp_2.RData")
```

```{r}
# Define time sequence
t <- seq(-1.6, 2, by = 0.1)

# Generate prediction dataset
x_pred <- lgpr::new_x(j_sub, t, group_by = NA, x = "wave_s")
```

```{r}
# Create function to wrangle the predictions to the correct format
process_predictions <- function(predicted) {
  # Extract and reshape predictions 
  x <- as.data.frame(predicted@x)
  y <- as.data.frame(predicted@h) %>%
    pivot_longer(cols = everything()) %>%
    dplyr::select(-name) %>%
    rename(y = value)

  # Combine 
  data <- cbind(x, y)
  
  return(data)
}


# Apply function for predictions for shared effect (f = fixed effect) and person-specific effect (r = random effects)
pred1 <-lgpr::pred(gp_2, x_pred,  verbose = FALSE)
gp_pred_f <- process_predictions(pred1)
pred2<- lgpr::pred(gp_2, verbose = FALSE)
gp_pred_r <- process_predictions(pred2)
```
#### Sort of works
```{r}
ggplot(gp_pred_f, aes(x = wave_s, y = y )) + 
  geom_smooth(color = project_colors[2], size = 1) +  
  geom_point(aes(y = y), color = project_colors[3], alpha = 0.7, size = 1.2)
```


```{r}
# Extract posterior samples from fit
posterior_samples <- pred(gp_2)

# Sort time points (months) for consistent plotting
order <- order(j_sub$wave_s)
waves <- j_sub$wave_s[order]
posterior_samples <- posterior_samples@f_comp$`gp_ns(wave_s)`[, order]

# Convert to tidy format for summarising
df_draws <- as_tibble(posterior_samples)
colnames(df_draws) <- as.character(seq_along(colnames(df_draws)))

# Reshape to long format
df_draws <- df_draws %>%
  mutate(draw = row_number()) %>%
  pivot_longer(
    cols = -draw,
    names_to = "wave_index",
    values_to = "value"
  ) %>%
  mutate(
    wave_index = as.integer(wave_index),
    wave = waves[wave_index]
  )

# Randomly select 1,000 posterior draws
set.seed(123)  # for reproducibility
draw_ids <- sample(unique(df_draws$draw), 100)

# Filter the data to keep only the selected draws
df_draws_sampled <- df_draws %>%
  filter(draw %in% draw_ids)

# Get original mean and SD of life satisfaction,
# to denormalise predictions back to the original scale of life satisfaction
mean_ls <- mean(j_data$cv_r1)
sd_ls <- sd(j_data$cv_r1)


# Compute 95% credible intervals per month
summary_df <- df_draws_sampled %>%
  group_by(wave) %>%
  summarise(
    lower_CI = quantile(value, 0.025),
    upper_CI = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(lower_CI = lower_CI * sd_ls + mean_ls,
         upper_CI = upper_CI * sd_ls + mean_ls
  )
```

```{r}
# For reproducibility
set.seed(123)

# Randomly sample 50 participants
rsample_ids <- sample(unique(j_sub$id), 3)

# Filter the data to include only the randomly selected participants
gp_pred_r_rsample <- gp_pred_r %>%
  filter(id %in% rsample_ids)
```
#### doesn't work for some reason
```{r}
ggplot()+
  geom_line(data = gp_pred_r_rsample, aes(waves, y, group = id), color = "grey70", linewidth = 0.4) +
  geom_ribbon(data = summary_df, aes(ymin = lower_CI, ymax = upper_CI, x = wave),
              alpha = 0.2, fill = "firebrick4") +
  geom_line(data = gp_pred_f, aes(x = waves, y = y),  color = "firebrick4", linewidth = 1) +
  ggtitle("Gaussian Process Regression") +
  theme(plot.title = element_text(size = 13, face = "bold"))
```

```{r}
plot_components(gp_2,pred = pred1, t_name = "wave_s")
```

```{r}
plot_components(gp_2, pred = pred2,  t_name = "wave_s")
```
###################################
```{r}

# simulate data using the mgcv package
dat <- mgcv::gamSim(1, n = 30, scale = 2)

# fit a simple GP model
fit1 <- brm(y ~ gp(x2), dat, chains = 2)
summary(fit1)
me1 <- conditional_effects(fit1, ndraws = 200, spaghetti = TRUE)
plot(me1, ask = FALSE, points = TRUE)

# fit a more complicated GP model
fit2 <- brm(y ~ gp(x0) + x1 + gp(x2) + x3, dat, chains = 2)
summary(fit2)
me2 <- conditional_effects(fit2, ndraws = 200, spaghetti = TRUE)
plot(me2, ask = FALSE, points = TRUE)

# fit a multivariate GP model
fit3 <- brm(y ~ gp(x1, x2), dat, chains = 2)
summary(fit3)
me3 <- conditional_effects(fit3, ndraws = 200, spaghetti = TRUE)
plot(me3, ask = FALSE, points = TRUE)

# compare model fit
loo(fit1, fit2, fit3)

# simulate data with a factor covariate
dat2 <- mgcv::gamSim(4, n = 90, scale = 2)

# fit separate gaussian processes for different levels of 'fac'
fit4 <- brm(y ~ gp(x2, by = fac), dat2, chains = 2)
summary(fit4)
plot(conditional_effects(fit4), points = TRUE)

save(fit1, file="gp_11.RData")
save(fit2, file="gp_12.RData")
save(fit3, file="gp_13.RData")
save(fit4, file="gp_14.RData")
```

