---
title: "gp_analysis"
output: html_document
date: "2026-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
########## Analysis for simulating GP with an artificial example
```{r}
#install.packages("tidyverse")
#install.packages("lgpr")
#install.packages("brms")
#install.packages("rstan")
#install.packages("rstanarm")
#install.packages("remotes")
#install.packages("tidybayes")
#install.packages("bridgesampling")
#install.packages("shinystan")
#install.packages("mgcv")
#remotes::install_github("michael-franke/aida-package")
#remotes::install_github("michael-franke/faintr")
#remotes::install_github("CogSciPrag/cspplot")

#install.packages("ggplot2")
#install.packages("gridExtra")

library(tidyverse)
library(dplyr)
library(lgpr)
library(brms)
library(rstan)
library(rstanarm)
library(remotes)
library(tidybayes)
library(bridgesampling)
library(shinystan)
library(mgcv)
library(aida)
library(faintr)
library(cspplot)
library(rstan)
library(plyr)
library(reshape2)
library(ggplot2)
library(gridExtra)

```
```{r}
# these options help Stan run faster
options(mc.cores = parallel::detectCores())

# use the CSP-theme for plotting
theme_set(theme_csp())

# global color scheme from CSP
project_colors = cspplot::list_colors() |> pull(hex)
# names(project_colors) <- cspplot::list_colors() |> pull(name)

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
}
```

```{r}
get_GP_simulation <- 
  function(x = seq(0,10, by = 0.1), 
           Intercept = 0, 
           slope = 1, 
           sigma = 1, 
           sigma_f=0.5, 
           lambda=100, 
           seed = NULL) {
    
    if (! is.null(seed)){
      set.seed(seed)
    }
    
    # number of points to generate prediction for
    N <- length(x)
    
    # linear predictor (vanilla LM)
    eta = Intercept + slope * x
    
    # kernel function (here: radial basis function kernel)
    get_covmatrix <- function(x, sigma_f, lambda) {
      K = matrix(0, nrow=N, ncol=N)
      for (i in 1:N) {
        for (j in 1:N) {
          K[i,j] = sigma_f^2 * exp(sqrt((i-j)^2) / (-2 *lambda))
        }
      }
      return(K)
    }
    
    # covariance matrix
    K <- get_covmatrix(x, sigma_f, lambda)
    
    # Gaussian process wiggles
    epsilon_gp <- mvtnorm::rmvnorm(
      n     = 1, 
      mean  = rep(0, N),
      sigma = K)[1,]
    
    # central tendency
    mu <- epsilon_gp + eta
    
    # data prediction
    y <- rnorm(N, mu, sigma)
      
    tibble(x, eta, mu, y) 
  }

plot_GP_simulation <- function(GP_simulation) {
  GP_simulation |> 
    ggplot(aes(x = x, y = eta )) + 
    geom_line(color = project_colors[1], size = 1.25) +
    geom_line(aes(y = mu), color = project_colors[2], size = 1.25) +
    geom_point(aes(y = y), color = project_colors[3], alpha = 0.7, size = 1.2)
}
```

```{r}
get_GP_simulation(
  x = seq(-1,1, length.out = 500), 
  Intercept = 0, 
  slope = 0.3, 
  sigma_f = 0.05, 
  lambda  = 20, 
  sigma = 0.05) |> 
  plot_GP_simulation()
```

```{r}
GP_simulation <- get_GP_simulation(
  x = seq(-1,1, length.out = 100), 
  Intercept = 0, 
  slope = 1, 
  sigma_f = 0.5, 
  lambda  = 2, 
  sigma = 0.0001,
  seed = 1996)

plot_GP_simulation(GP_simulation)
```

```{r}
# results: hide
fit_GPR <- 
  brms::brm(
    formula = y ~ gp(x) + x,
    data    = GP_simulation,
    iter    = 2000
  )
```

```{r}
summary(fit_GPR)
```

```{r}
conditional_effects(fit_GPR)
```

########### Artificial data fits fine
############ Data of WHR for Benevolence in Netherlands
```{r}
x <- c(2007, 2008, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024)
y <- c(74.9, 77.1, 75.3, 74.0, 69.2, 70.8, 73.5, 66.6, 64.5, 65.8, 57.1, 62.3, 55.8, 68.1, 64.1, 64.0, 60.7)

data<- data.frame(x, y)

ggplot(data, aes(x = x, y = y )) + 
    geom_line(color = project_colors[1], size = 1.25) +
    geom_point(aes(y = y), color = project_colors[3], alpha = 0.7, size = 1.2)
```

```{r}
data$x_s <- (scale(data$x))[,1]

fit_GPR_2 <- brms::brm(
  y ~ gp(x_s, kappa = 1e-8),
  data = data,
  iter = 4000
)


summary(fit_GPR_2)
```

########### 17 observations are not enough

```{r}
newdata <- data.frame(
  x_s = seq(min(data$x_s), max(data$x_s), length.out = 200)
)

pred <- fitted(
  fit_GPR_2,
  newdata = newdata,
  re_formula = NA
)

```

############################################### Analysis of Japan panel on covid, 30 waves

```{r}
panel_japan <- read.csv("data.csv", stringsAsFactors = FALSE)

j_data<- panel_japan[, c("cw_ID", "cv_r1", "wave", "sleep_x", "only_jpn")]

j_data <- j_data %>%
  mutate(
    wave_s   = as.numeric(scale(wave)),
    #id       = as.integer(factor(cw_ID)),
    id       = as.factor(factor(cw_ID)),
    cv_r1    = as.numeric(scale(cv_r1)),
    sleep_x  = as.numeric(scale(sleep_x)),
    only_jpn = as.numeric(scale(as.numeric(only_jpn))
  ))

j_data<- j_data[, c("id", "cv_r1", "wave_s")]

head(j_data)
```

# BRMS analysis on just one person - 30 observations, does not work
```{r}
j_data_1 <- subset(j_data, id == "8491")

ggplot(j_data_1, aes(x = wave_s, y = cv_r1 )) + 
    geom_line(color = project_colors[1], size = 1.25) +
  geom_smooth(color = project_colors[2], size = 1) +  
  geom_point(aes(y = cv_r1), color = project_colors[3], alpha = 0.7, size = 1.2)

ggplot(j_data_1, aes(x = wave_s, y = sleep_x )) + 
    geom_line(color = project_colors[1], size = 1.25) +
    geom_smooth(color = project_colors[2], size = 1) + 
    geom_point(aes(y = sleep_x), color = project_colors[3], alpha = 0.7, size = 1.2)

ggplot(j_data_1, aes(x = wave_s, y = only_jpn )) + 
    geom_smooth(color = project_colors[2], size = 1) +
    geom_line(color = project_colors[1], size = 1.25) +
  geom_point(aes(y = only_jpn), color = project_colors[3], alpha = 0.7, size = 1.2)
```
##### just one person, 17 obs - not working
```{r}
fit_GPR_3 <- brms::brm(
  cv_r1 ~ gp(wave_s, kappa = 1e-2),
  data = j_data_1,
  iter = 4000
)

summary(fit_GPR_3)

newdata <- data.frame(
  wave_s = seq(min(j_data_1$wave_s), max(j_data_1$wave_s), length.out = 200)
)

pred <- fitted(
  fit_GPR_3,
  newdata = newdata,
  re_formula = NA
)
```
############################### Analysis with lgpr to study several people
```{r}
# For reproducibility
set.seed(123)

ggplot() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(x = "wave", y = "Dread") +
  theme(legend.position = "none") +
  theme(aspect.ratio=0.9) +
geom_smooth(data = j_data, aes(x=wave_s, y=cv_r1)) +
  ggtitle("Average dreadfulness levels") +
  theme(plot.title = element_text(size = 13, face = "bold")) 
```

```{r}
# For reproducibility
set.seed(123)

# Define mean and SD of the log-normal distribution
mean_priordistribution_lengthscale <- 0
sd_priordistribution_lengthscale <- 1

# Define warping function
warpfun <- function(time,warpingparameter){
  2 / (1 + exp( - warpingparameter * time)) - 1
}

# Define kernel with warping function
kernelGP <- function(X1, X2, lengthscale, alpha, warpingparameter, fun=function(x,w){x}){
  
  X1mat <- matrix(rep(fun(X1,warpingparameter),length(X2)),nrow=length(X1))
  X2mat <- matrix(rep(fun(X2,warpingparameter),length(X1)),nrow=length(X1),byrow=TRUE)
  
  alpha**2 * exp(- .5 * (X1mat - X2mat) * (X1mat - X2mat) / lengthscale**2 )
  
}

n <- 100 
X <- seq(-100, 100, length=n) 
mu = rep(0,length(X))

# Set up a 4x4 plotting panel
par(mfrow = c(4, 4), mar = c(2, 2, 2, 2))

# Loop to generate 16 plots
for (i in 1:16) {

# Lengthscale parameter
lengthscale1 <- dlnorm(1, meanlog = mean_priordistribution_lengthscale, sdlog = sd_priordistribution_lengthscale)

# Magnitude parameter (potentially modify prior distribution here!)
alpha1 <- dlnorm(1, meanlog = 0, sdlog = 1)

# Warping parameter (potentially modify prior distribution here!)
warpingparameter <- dlnorm(1, meanlog=-0.7,sdlog=0.9)
covm = kernelGP(X, X, lengthscale = lengthscale1, alpha = alpha1, warpingparameter = warpingparameter, fun = warpfun)

# Intercept
beta0 <- 0

  # Draw a line from the GP
  draw_f <- mvtnorm::rmvnorm(1, mean = mu, sigma = covm)
  
  # Plot the line
  plot(X, beta0 + draw_f, type = "l", col = "grey", lwd = 2,
       ylim = c(beta0 - 2 * alpha1, beta0 + 2 * alpha1),
       ylab = "y", xlab = "x", main = paste("Plot", i))
}

prior <- list(
  alpha = log_normal(0, 1), 
  ell = log_normal(0, 1),
  wrp = log_normal(-0.7, 0.9)
)
```


```{r}
j_sub <- subset(j_data, id %in% c("42040", "8491", "15569", "94338"))

#gp <- lgp(
#  cv_r1 ~
#    gp_ns(wave_s) +  # population-level GP
#    gp_ns(wave_s)*zs(id),# group-level GP
#  data = j_sub,
#  sample_f = TRUE
#)

gp_2 <- lgp(
  cv_r1 ~
    gp_ns(wave_s) +  zs(id),
  data = j_sub,
  sample_f = TRUE
)

summary(gp_2)

print(gp_2)
save(gp_2, file="gp_2.RData")
load("gp_2.RData")
```

```{r}
# Define time sequence
t <- seq(-1.6, 2, by = 0.1)

# Generate prediction dataset
x_pred <- lgpr::new_x(j_sub, t, group_by = NA, x = "wave_s")
```

```{r}
# Create function to wrangle the predictions to the correct format
process_predictions <- function(predicted) {
  # Extract and reshape predictions 
  x <- as.data.frame(predicted@x)
  y <- as.data.frame(predicted@h) %>%
    pivot_longer(cols = everything()) %>%
    dplyr::select(-name) #%>%   rename(y = value)

  # Combine 
  data <- cbind(x, y)
  
  return(data)
}


# Apply function for predictions for shared effect (f = fixed effect) and person-specific effect (r = random effects)
pred1 <-lgpr::pred(gp_2, x_pred,  verbose = FALSE)
gp_pred_f <- process_predictions(pred1)
pred2<- lgpr::pred(gp_2, verbose = FALSE)
gp_pred_r <- process_predictions(pred2)
```
#### Sort of works
```{r}
ggplot(gp_pred_f, aes(x = wave_s, y = value )) + 
  geom_smooth(color = project_colors[2], size = 1) +  
  geom_point(aes(y = value), color = project_colors[3], alpha = 0.7, size = 1.2)
```


```{r}
# Extract posterior samples from fit
posterior_samples <- pred(gp_2)

# Sort time points (months) for consistent plotting
order <- order(j_sub$wave_s)
waves <- j_sub$wave_s[order]
posterior_samples <- posterior_samples@f_comp$`gp_ns(wave_s)`[, order]

# Convert to tidy format for summarising
df_draws <- as_tibble(posterior_samples)
colnames(df_draws) <- as.character(seq_along(colnames(df_draws)))

# Reshape to long format
df_draws <- df_draws %>%
  mutate(draw = row_number()) %>%
  pivot_longer(
    cols = -draw,
    names_to = "wave_index",
    values_to = "value"
  ) %>%
  mutate(
    wave_index = as.integer(wave_index),
    wave = waves[wave_index]
  )

# Randomly select 1,000 posterior draws
set.seed(123)  # for reproducibility
draw_ids <- sample(unique(df_draws$draw), 100)

# Filter the data to keep only the selected draws
df_draws_sampled <- df_draws %>%
  filter(draw %in% draw_ids)

# Get original mean and SD of life satisfaction,
# to denormalise predictions back to the original scale of life satisfaction
mean_ls <- mean(j_data$cv_r1)
sd_ls <- sd(j_data$cv_r1)


# Compute 95% credible intervals per month
summary_df <- df_draws_sampled %>%
  group_by(wave) %>%
  summarise(
    lower_CI = quantile(value, 0.025),
    upper_CI = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(lower_CI = lower_CI * sd_ls + mean_ls,
         upper_CI = upper_CI * sd_ls + mean_ls
  )
```

```{r}
# For reproducibility
set.seed(123)

# Randomly sample 50 participants
rsample_ids <- sample(unique(j_sub$id), 3)

# Filter the data to include only the randomly selected participants
gp_pred_r_rsample <- gp_pred_r %>%
  filter(id %in% rsample_ids)
```
#### doesn't work for some reason
```{r}
ggplot()+
  geom_line(data = gp_pred_r_rsample, aes(waves, y, group = id), color = "grey70", linewidth = 0.4) +
  geom_ribbon(data = summary_df, aes(ymin = lower_CI, ymax = upper_CI, x = wave),
              alpha = 0.2, fill = "firebrick4") +
  geom_line(data = gp_pred_f, aes(x = waves, y = y),  color = "firebrick4", linewidth = 1) +
  ggtitle("Gaussian Process Regression") +
  theme(plot.title = element_text(size = 13, face = "bold"))
```

```{r}
plot_components(gp_2,pred = pred1, t_name = "wave_s")
```

```{r}
plot_components(gp_2, pred = pred2,  t_name = "wave_s")
```
###################################
```{r}

# simulate data using the mgcv package
dat <- mgcv::gamSim(1, n = 30, scale = 2)

# fit a simple GP model
fit1 <- brm(y ~ gp(x2), dat, chains = 2)
summary(fit1)
me1 <- conditional_effects(fit1, ndraws = 200, spaghetti = TRUE)
plot(me1, ask = FALSE, points = TRUE)

# fit a more complicated GP model
fit2 <- brm(y ~ gp(x0) + x1 + gp(x2) + x3, dat, chains = 2)
summary(fit2)
me2 <- conditional_effects(fit2, ndraws = 200, spaghetti = TRUE)
plot(me2, ask = FALSE, points = TRUE)

# fit a multivariate GP model
fit3 <- brm(y ~ gp(x1, x2), dat, chains = 2)
summary(fit3)
me3 <- conditional_effects(fit3, ndraws = 200, spaghetti = TRUE)
plot(me3, ask = FALSE, points = TRUE)

# compare model fit
loo(fit1, fit2, fit3)

# simulate data with a factor covariate
dat2 <- mgcv::gamSim(4, n = 90, scale = 2)

# fit separate gaussian processes for different levels of 'fac'
fit4 <- brm(y ~ gp(x2, by = fac), dat2, chains = 2)
summary(fit4)
plot(conditional_effects(fit4), points = TRUE)

save(fit1, file="gp_11.RData")
save(fit2, file="gp_12.RData")
save(fit3, file="gp_13.RData")
save(fit4, file="gp_14.RData")

load("gp_11.RData")
load("gp_12.RData")
load("gp_13.RData")
load("gp_14.RData")
```

################################################demo from https://gist.github.com/jkeirstead/6271854

```{r}
## 1. Simulate a process with no data
## The very small sigma_sq value is necessary to avoid an error.  Don't set it to zero.
x <- seq(-5, 5, 0.2)
n <- length(x)
fit <- stan(file="gp-sim.stan", data=list(x=x, N=n, eta_sq=1, rho_sq=0.5, sigma_sq=0.0001),
            iter=200, chains=3)
sims <- extract(fit, permuted=TRUE)

## Rearrange the data and plot it
data <- adply(sims$y, 2)
tmp <- melt(data)
names(tmp) <- c("xid", "group", "y")
tmp <- mutate(tmp, x=x[xid])
fig2a <- ggplot(tmp, aes(x=x, y=y)) +
  geom_line(aes(group=group), colour="#999999", alpha=0.3) +
  theme_bw()
```
```{r}
## 2. Simulate with a few noise-free data points.
## Again pretend the noise is almost zero, but not quite.
x1 <- c(-4, -3, -1, 0, 2)
y1 <- c(-2, 0, 1, 2, -1)
            
## Parameter value fixed in given example
fit <- stan(file="gp-predict.stan", data=list(x1=x1, y1=y1, N1=length(x1), 
                                      x2=x, N2=length(x), eta_sq=1, rho_sq=0.5, sigma_sq=0.0001),
            iter=200, chains=3)
sims <- extract(fit, permuted=TRUE)

## Rearrange the data and plot it
data <- adply(sims$y, 2)
tmp <- melt(data)
names(tmp) <- c("xid", "group", "y")
tmp <- mutate(tmp, x=x[xid])
fig2b <- ggplot(tmp, aes(x=x, y=y)) +
  geom_line(aes(group=group), colour="#999999", alpha=0.3) +
  theme_bw() +
  geom_point(data=data.frame(x=x1, y=y1))

```
```{r}

## 3.  Adding more noise is easy.  Just change sigma_sq
sigma.n <- 0.1
fit <- stan(file="gp-predict.stan", data=list(x1=x1, y1=y1, N1=length(x1), 
                                      x2=x, N2=length(x), eta_sq=1, rho_sq=1, sigma_sq=sigma.n^2),
            iter=200, chains=3)
sims <- extract(fit, permuted=TRUE)

## Rearrange the data and plot
data <- adply(sims$y, 2)
tmp <- melt(data)
names(tmp) <- c("xid", "group", "y")
tmp <- mutate(tmp, x=x[xid])
fig2c <- ggplot(tmp, aes(x=x, y=y)) +
  geom_line(aes(group=group), colour="#999999", alpha=0.3) +
  theme_bw() +
  geom_point(data=data.frame(x=x1, y=y1)) +
  geom_errorbar(data=data.frame(x=x1, y=y1), aes(x=x,y=NULL,ymin=y-2*sigma.n, ymax=y+2*sigma.n), width=0.2) 

## Save plots for the web
w <- 6
h <- 4
ggsave("fig2a-rstan.png", fig2a, width=w, height=h, dpi=150)
ggsave("fig2b-rstan.png", fig2b, width=w, height=h, dpi=150)
ggsave("fig2c-rstan.png", fig2c, width=w, height=h, dpi=150)
grid.arrange(fig2a, fig2b, fig2c, ncol = 3)
```

####################################### Analysis from https://github.com/MaggieLieu/STAN_tutorials/tree/master/Gaussian_processes_I

```{r}
N=100

data = list(
  N = N,
  x = seq(-10, 10, length.out=N), #equally spaced points
  alpha = 1,
  rho = 1
)

ndraws = 50 # number of draws
fit = stan(file='sim_gp.stan', data=data, algorithm ='Fixed_param',warmup = 0, chains = 1, iter=ndraws)
params = extract(fit)
xr=c(-10,10)
yr=c(-5,5)
plot(xr, yr, ty='n')
for(i in 1:ndraws){
  lines(data$x, params$f[i,], pch=20, col=rgb(0,0.55,0.5, 0.2), cex=0.5)
}
points(data$x, params$f[ndraws,], pch=20, col='black', cex=1)
lines(data$x, colMeans(params$f),lwd=3)
```

```{r}
yconf = sapply(1:N, function(x) quantile(params$f[,x], probs=c(0.003,0.05,0.32,0.68,0.95,0.997)))

plot(xr,yr, ty='n', xlab='t', ylab='y')
for(i in 1:3){
  polygon(x=c(data$x, rev(data$x)), y=c(yconf[i,], rev(yconf[i+3,])), border=NA, col=rgb(0,0,0,0.1))
}
lines(data$x, colMeans(params$f),lwd=3)
```

```{r}
lcdata = read.csv('kplr003632418-2009131105131_llc.csv')
plot(lcdata$TIME, lcdata$SAP_FLUX, pch=20, xlab='Time [days]', ylab='Flux [e-/s]', cex=0.5)
bad = which(is.na(lcdata$SAP_FLUX)) #remove missing data
x = lcdata$TIME[-bad] 
y = (lcdata$SAP_FLUX[-bad] - mean(lcdata$SAP_FLUX[-bad]))/sd(lcdata$SAP_FLUX[-bad]) #Normalise flux to have mean 0 and sd 1
plot(x,y, pch=20, cex=0.5)
data = list(
  N = length(x),
  x = x,
  y = y
)
fit = stan(file='method2.stan', data=data,chains = 1, iter=2000)
params = extract(fit)
pairs(fit)
mean(params$rho)
mean(params$alpha)
mean(params$sigma)
```

```{r}
N =500
gen_data = list(
  N = N,
  x = seq(120, 132, length.out=N), #equally spaced points
  alpha = mean(params$alpha),
  rho = mean(params$rho), 
  sigma = mean(params$sigma)
)

ndraws = 50 # number of draws
gen_sam = stan(file = 'sim_gp.stan', data=gen_data, algorithm ='Fixed_param',warmup = 0, chains = 1, iter=ndraws)
gen_params = extract(gen_sam)
xr = c(120,132)
yr = c(-4,4)
plot(xr, yr, ty='n')
for(i in 1:ndraws){
  lines(gen_data$x, gen_params$f[i,], pch=20, col=rgb(0,0.55,0.5, 0.2))
}
points(data$x, data$y, pch=20, col='orange', cex=0.5)
legend('topright', y = c('predicted','data'), col = c(rgb(0,0.55,0.5, 0.2),'orange'), bty='n', pch=c(NA,20), lty=c(1,NA))
```

```{r}
N_predict = 100
x_predict = seq(range(data$x)[1], 132, length.out=N_predict)
N_obs = 200
Obs = sort(sample(1:length(x), N_obs))
pred_data <- list(N1=N_obs, x1=x[Obs], y1=y[Obs], N2=N_predict, x2=x_predict)

pred_fit = stan(file='method1.stan', data =pred_data, iter=2000, chains=1)
pred_params = extract(pred_fit)
yr = c(-2,2)
xr = c(120, 132)
plot(xr, yr, ty='n')
for( i in 1:100){
  lines(x_predict, pred_params$f[i,(N_obs+1):(N_obs+N_predict)], col=rgb(0,0,0,0.1))
}
points(data$x, data$y, pch=20, col='orange', cex=0.3)
```

```{r}
N_predict = 100
x_predict = seq(range(x)[1], 132, length.out=N_predict)
N_obs = 3
Obs = sort(sample(1:length(x), N_obs))
pred_data <- list(rho=mean(params$rho),alpha=mean(params$alpha), sigma=mean(params$sigma), N=N_obs, x=x[Obs], y=y[Obs], N_predict=N_predict, x_predict=x_predict)

pred_opt_fit <- stan(file='predict_gauss.stan', data=pred_data, iter=1000, warmup=0,
                     chains=1, algorithm="Fixed_param")
pred_params = extract(pred_opt_fit)
yr = c(-2,2)
xr = c(120, 132)
plot(xr, yr, ty='n')
for( i in 1:100){
  lines(x_predict, pred_params$f_predict[i,], col=rgb(0,0,0,0.1))
}
points(x, y, pch=20, col='orange', cex=0.3)
points(pred_data$x, pred_data$y, pch=20, col='orange', cex=2)
points(pred_data$x, pred_data$y, cex=2)
```

```{r}
yconf = sapply(1:N_predict, function(x) quantile(pred_params$f[,x], probs=c(0.003,0.05,0.32,0.68,0.95,0.997)))

plot(xr,yr, ty='n', xlab='t', ylab='y')
for(i in 1:3){
  polygon(x=c(x_predict, rev(x_predict)), y=c(yconf[i,], rev(yconf[i+3,])), border=NA, col=rgb(0,0,0,0.1))
}
lines(x_predict, colMeans(pred_params$f),lwd=3)
```

